### 2018-08-26 doc/Paper.tex
Update PDF to bbc110c18bcab415f0b5b60cda790011de514098

\documentclass[runningheads]{llncs}

\usepackage{amsmath}
\usepackage{color}
\usepackage{comment}
\usepackage{float}
\usepackage{graphicx}
\usepackage{url}
\usepackage{subcaption}
\usepackage{wrapfig}

\captionsetup{compatibility=false}

%% ODER: format ==         = "\mathrel{==}"
%% ODER: format /=         = "\neq "
%
%
\makeatletter
\@ifundefined{lhs2tex.lhs2tex.sty.read}%
  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
   \newcommand\SkipToFmtEnd{}%
   \newcommand\EndFmtInput{}%
   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
  }\SkipToFmtEnd

\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{stmaryrd}
\DeclareFontFamily{OT1}{cmtex}{}
\DeclareFontShape{OT1}{cmtex}{m}{n}
  {<5><6><7><8>cmtex8
   <9>cmtex9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
\DeclareFontShape{OT1}{cmtex}{m}{it}
  {<-> ssub * cmtt/m/it}{}
\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
\DeclareFontShape{OT1}{cmtt}{bx}{n}
  {<5><6><7><8>cmtt8
   <9>cmbtt9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
\DeclareFontShape{OT1}{cmtex}{bx}{n}
  {<-> ssub * cmtt/bx/n}{}
\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU

\newcommand{\Sp}{\hskip.33334em\relax}

\newcommand{\Conid}[1]{\mathit{#1}}
\newcommand{\Varid}[1]{\mathit{#1}}
\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
\newcommand{\plus}{\mathbin{+\!\!\!+}}
\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
\newcommand{\sequ}{\mathbin{>\!\!\!>}}
\renewcommand{\leq}{\leqslant}
\renewcommand{\geq}{\geqslant}
\usepackage{polytable}

%mathindent has to be defined
\@ifundefined{mathindent}%
  {\newdimen\mathindent\mathindent\leftmargini}%
  {}%

\def\resethooks{%
  \global\let\SaveRestoreHook\empty
  \global\let\ColumnHook\empty}
\newcommand*{\savecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
\newcommand*{\restorecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
\newcommand*{\aligncolumn}[2]%
  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}

\resethooks

\newcommand{\onelinecommentchars}{\quad-{}- }
\newcommand{\commentbeginchars}{\enskip\{-}
\newcommand{\commentendchars}{-\}\enskip}

\newcommand{\visiblecomments}{%
  \let\onelinecomment=\onelinecommentchars
  \let\commentbegin=\commentbeginchars
  \let\commentend=\commentendchars}

\newcommand{\invisiblecomments}{%
  \let\onelinecomment=\empty
  \let\commentbegin=\empty
  \let\commentend=\empty}

\visiblecomments

\newlength{\blanklineskip}
\setlength{\blanklineskip}{0.66084ex}

\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
\let\hspre\empty
\let\hspost\empty
\newcommand{\NB}{\textbf{NB}}
\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}

\EndFmtInput
\makeatother
%
%
%
%
%
%
% This package provides two environments suitable to take the place
% of hscode, called "plainhscode" and "arrayhscode". 
%
% The plain environment surrounds each code block by vertical space,
% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
% similar to formulas. Note that if these dimensions are changed,
% the spacing around displayed math formulas changes as well.
% All code is indented using \leftskip.
%
% Changed 19.08.2004 to reflect changes in colorcode. Should work with
% CodeGroup.sty.
%
\ReadOnlyOnce{polycode.fmt}%
\makeatletter

\newcommand{\hsnewpar}[1]%
  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}

% can be used, for instance, to redefine the code size, by setting the
% command to \small or something alike
\newcommand{\hscodestyle}{}

% The command \sethscode can be used to switch the code formatting
% behaviour by mapping the hscode environment in the subst directive
% to a new LaTeX environment.

\newcommand{\sethscode}[1]%
  {\expandafter\let\expandafter\hscode\csname #1\endcsname
   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}

% "compatibility" mode restores the non-polycode.fmt layout.

\newenvironment{compathscode}%
  {\par\noindent
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed\)%
   \par\noindent
   \ignorespacesafterend}

\newcommand{\compaths}{\sethscode{compathscode}}

% "plain" mode is the proposed default.
% It should now work with \centering.
% This required some changes. The old version
% is still available for reference as oldplainhscode.

\newenvironment{plainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newenvironment{oldplainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

% Here, we make plainhscode the default environment.

\newcommand{\plainhs}{\sethscode{plainhscode}}
\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
\plainhs

% The arrayhscode is like plain, but makes use of polytable's
% parray environment which disallows page breaks in code blocks.

\newenvironment{arrayhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\parray}%
  {\endparray\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newcommand{\arrayhs}{\sethscode{arrayhscode}}

% The mathhscode environment also makes use of polytable's parray 
% environment. It is supposed to be used only inside math mode 
% (I used it to typeset the type rules in my thesis).

\newenvironment{mathhscode}%
  {\parray}{\endparray}

\newcommand{\mathhs}{\sethscode{mathhscode}}

% texths is similar to mathhs, but works in text mode.

\newenvironment{texthscode}%
  {\(\parray}{\endparray\)}

\newcommand{\texths}{\sethscode{texthscode}}

% The framed environment places code in a framed box.

\def\codeframewidth{\arrayrulewidth}
\RequirePackage{calc}

\newenvironment{framedhscode}%
  {\parskip=\abovedisplayskip\par\noindent
   \hscodestyle
   \arrayrulewidth=\codeframewidth
   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
   \hline\framedhslinecorrect\\{-1.5ex}%
   \let\endoflinesave=\\
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \framedhslinecorrect\endoflinesave{.5ex}\hline
   \endtabular
   \parskip=\belowdisplayskip\par\noindent
   \ignorespacesafterend}

\newcommand{\framedhslinecorrect}[2]%
  {#1[#2]}

\newcommand{\framedhs}{\sethscode{framedhscode}}

% The inlinehscode environment is an experimental environment
% that can be used to typeset displayed code inline.

\newenvironment{inlinehscode}%
  {\(\def\column##1##2{}%
   \let\>\undefined\let\<\undefined\let\\\undefined
   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
   \def\fromto##1##2##3{##3}%
   \def\nextline{}}{\) }%

\newcommand{\inlinehs}{\sethscode{inlinehscode}}

% The joincode environment is a separate environment that
% can be used to surround and thereby connect multiple code
% blocks.

\newenvironment{joincode}%
  {\let\orighscode=\hscode
   \let\origendhscode=\endhscode
   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
   %\let\SaveRestoreHook=\empty
   %\let\ColumnHook=\empty
   %\let\resethooks=\empty
   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
  {\origendhscode
   \global\let\hscode=\orighscode
   \global\let\endhscode=\origendhscode}%

\makeatother
\EndFmtInput
%

%%format .*.       = "\mathbin{.\!\!*\!\!.}"
%%format .=.       = "\mathbin{.\!\!=\!\!.}"
%%format .<.       = "\mathbin{.\!\!<\!\!.}"
%%format sp = " "
%%format qq = "''\!"
%%format .$$ = ".\$_m"
%%format parent' = "parent_m"
%%format  <*>       = "\mathbin{\text{\small\ttfamily{<*>}}}"
%%format  <$>       = "\mathbin{\text{\small\ttfamily{<\$>}}}"
%%format t1 = "t_1"
%%format t2 = "t_2"
%%format constructor' = "constructor_m"
%%%%format Leaf = "{\sf Leaf}"
%%%%format Fork = "{\sf Fork}"
%%format Fork3 = "{\sf Fork3}"
%%%%%%format Root = "{\sf Root}"
%%format CRoot = "C_{\mathit{Root}}"
%%%%%%"{\sf C_{Root}}"
%%format CLeaf = "C_{\mathit{Leaf}}"
%%%%%"{\sf C_{Leaf}}"
%%format CFork = "C_{\mathit{Fork}}"
%%%%% "{\sf C_{Fork}}"
%%format lexeme_Leaf = "lexeme_{\sf Leaf}"
%%format MemoRoot = "{\sf Memo_{Root}}"
%%format MemoFork = "\mathit{Fork}_m"
%%%%%%"{\sf Fork}_m"
%%%%%% "{\sf Memo_{Fork}}"
%%format MemoLeaf = "\mathit{Leaf}_m"
%%%%%%"{\sf Leaf}_m"
%%%%%% "{\sf Memo_{Leaf}}"
%%format treeLookup = "lookup_{MT}"
%%%format MemoTable = "{Cache}"
%%format buildMTree = "\mathit{build}_m"
%%%%%%"build_{MT}"
%%format constructor_m = "\mathit{constructor}_m"
%%format tree_m = "\mathit{tree}_m"
%%format left_m = "\mathit{left}_m"
%%format right_m = "\mathit{right}_m"
%%format constructorM_m = "\mathit{constructorM}_m"
%%format treeM_m = "\mathit{treeM}_m"
%%format leftM_m = "\mathit{leftM}_m"
%%format rightM_m = "\mathit{rightM}_m"
%%format up_m = "\mathit{up}_m"
%%format down_m = "\mathit{down}_m"
%%format modify_m = "\mathit{modify}_m"
%%format mkAG_m = "mkAG_m"
%%format Cxt_m = "\mathit{Cxt}_m"
%%format Root_m = "Root_m"
%%format Top_m = "Top_m"
%%format L_m = "L_m"
%%format R_m = "R_m"
%%format MemoTree = "\mathit{Tree}_m"
%%format ZipperMemoTree = "\mathit{Zipper}_m"
%%format MemoAGTree = "\mathit{AGTree}_m"
%%format C_Memo_RootLet    = "C_{\mathit{Memo\_RootLet   }}"
%%format C_Memo_Let        = "C_{\mathit{Memo\_Let       }}"
%%format C_Memo_In         = "C_{\mathit{Memo\_In        }}"
%%format C_Memo_ConsAssign = "C_{\mathit{Memo\_ConsAssign}}"
%%format C_Memo_ConsLet    = "C_{\mathit{Memo\_ConsLet   }}"
%%format C_Memo_EmptyList  = "C_{\mathit{Memo\_EmptyList }}"
%%format C_Memo_Plus       = "C_{\mathit{Memo\_Plus      }}"
%%format C_Memo_Variable   = "C_{\mathit{Memo\_Variable  }}"
%%format C_Memo_Constant   = "C_{\mathit{Memo\_Constant  }}"
%%format lexemeConsAssign  = "lexeme_{\mathit{ConsAssign}}"
%%format lexemmeConsLet    = "lexeme_{\mathit{ConsLet}}"
%%format calculateMemo     = "calculate_{\mathit{Memo}}"
%%format errsAlgolMemo     = "errs_{\mathit{AlgolMemo}}"
%%format errorsMemo             = "errors_{\mathit{Memo}}"
%%format algol68                = "algol_{\mathit{68}}"
%%format ata_algol68            = "ata\_algol_{\mathit{68}}"
%%format LetSem.constructorMemo = "LetSem.constructor_{\mathit{Memo}}"
%%format lexemeString           = "lexeme_{\mathit{String}}"
%%format upGetVarValueMemo      = "up_{\mathit{GetVarValueMemo}}"
%%format lexemeInt              = "lexeme_{\mathit{Int}}"
%%format Child1              = "Child_{\mathit{1}}"
%%format Child2              = "Child_{\mathit{2}}"
%%format Child3              = "Child_{\mathit{3}}"
%%format Childi              = "Child_{\mathit{i}}"
%%format constructorMemo     = "constructor_{\mathit{Memo}}"
%%format at = "@"
%%format Algol68m               = "Algol68_{\mathit{m}}"

% Formatting
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% A command for declaring todos
\newcommand{\TODO}[1]{{\color[rgb]{1,0,0}\textbf{TODO:}\textit{#1}}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\TODO{What's the title?}}

% TODO(twesterhout): This looks ugly... Someone, reformat it, please :)
\author{Jo{\~a}o Paulo Fernandes\inst{1}%
   \and Pedro Martins\inst{2}%
   \and Alberto Pardo\inst{3}%
   \and Jo{\~a}o Saraiva\inst{4}%
   \and Marcos Viera\inst{3}%
   \and Tom Westerhout\inst{5}%
}
\institute{%
  LISP/Release -- Universidade da Beira Interior, Portugal \email{jpf@di.ubi.pt} \and
  University of California, Irvine, USA \email{pribeiro@uci.edu} \and
  Universidad de la  Rep\'{u}blica, Uruguay \email{\{pardo,mviera\}@fing.edu.uy} \and
  Universidade do Minho, Portugal \email{saraiva@di.uminho.pt} \and
  Radboud University, The Netherlands \email{twesterhout@student.ru.nl}%
}

\date{}

\maketitle

\begin{abstract}
  \TODO{What's the abstract?}

\keywords{%
       Embedded Domain Specific Languages
  \and Zipper data structure
  \and Memoization
  \and Attribute Grammars
  \and Higher-Order Attribute Grammars
  \and Functional Programming%
}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:introduction}

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[3]{}\Varid{main}\mathbin{::}\Conid{IO}\;(){}\<[E]%
\\
\>[3]{}\Varid{main}\mathrel{=}\Varid{putStrLn}\;\text{\tt \char34 Hello~world!\char34}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Functional Zippers}
  The zipper data structure was originally conceived by
  Huet\cite{huet1997zipper} to solve the problem of representing a tree together
  with a subtree that is the focus of attention, where that focus may move left,
  right, up or down the tree. Bla-bla-bla...

  Application to binary trees...
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[3]{}\mathbf{data}\;\Conid{Tree}\;\Varid{a}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mathrel{=}\Conid{Fork}\;(\Conid{Tree}\;\Varid{a})\;(\Conid{Tree}\;\Varid{a}){}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mid \Conid{Leaf}\,!\Varid{a}{}\<[E]%
\\[\blanklineskip]%
\>[3]{}\mathbf{data}\;\Conid{Path}\;\Varid{a}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mathrel{=}\Conid{Top}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mid \Conid{Left}\,!(\Conid{Path}\;\Varid{a})\;(\Conid{Tree}\;\Varid{a}){}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mid \Conid{TreeRight}\;(\Conid{Tree}\;\Varid{a})\,!(\Conid{Path}\;\Varid{a}){}\<[E]%
\\[\blanklineskip]%
\>[3]{}\mathbf{data}\;\Conid{Zipper}\;\Varid{a}\mathrel{=}\Conid{Zipper}\,!(\Conid{Path}\;\Varid{a})\;(\Conid{Tree}\;\Varid{a}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

  Application to lists...
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[3]{}\mathbf{data}\;\Conid{Path}\;\Varid{a}\mathrel{=}\Conid{Path}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
\\[\blanklineskip]%
\>[3]{}\mathbf{data}\;\Conid{Zipper}\;\Varid{a}\mathrel{=}\Conid{Zipper}\,!(\Conid{Path}\;\Varid{a})\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

  Generic zipper...

  An application of generic zipper that we will consider is embedding of
  attribute grammars.

\section{Attribute Grammars}
  What attribute grammars are...

  Repmin as two traversals...

  Repmin as a circular program...

  Repmin as an AG...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec:related-work}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}\label{sec:conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{References}
  % As per the LLNCS guidelines
  \bibliographystyle{splncs04}
  \bibliography{References}

\end{document}

### 2018-08-29 doc/Paper.tex
Update PDF to 772f0f944bf2079b09c69ccd4afe019641c0ca7f

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Functional Zippers}
+  Zipper is a data structure commonly used in functional programming for
+  traversal with fast local updates. The zipper data structure was originally
+  conceived by Huet\cite{huet1997zipper} in the context of trees. We will,
+  however, first consider a simpler problem: a bidirectional list traversal.
+
+  Suppose that we would like to update a list at a specific position:
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{19}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\Varid{modify}\mathbin{::}(\Varid{a}\to [\mskip1.5mu \Varid{a}\mskip1.5mu])\to \Conid{Int}\to [\mskip1.5mu \Varid{a}\mskip1.5mu]\to [\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
+\\
+\>[3]{}\Varid{modify}\;\Varid{f}\;\Varid{i}\;\Varid{xs}\mathrel{=}\Varid{helper}\;[\mskip1.5mu \mskip1.5mu]\;\Varid{xs}\;\mathrm{0}{}\<[E]%
+\\
+\>[3]{}\hsindent{1}{}\<[4]%
+\>[4]{}\mathbf{where}{}\<[E]%
+\\
+\>[4]{}\hsindent{1}{}\<[5]%
+\>[5]{}\Varid{helper}\;\Varid{before}\;(\Varid{x}\mathbin{:}\Varid{after})\,!\Varid{j}{}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{}\mid \Varid{j}\equiv \Varid{i}{}\<[19]%
+\>[19]{}\mathrel{=}\Varid{before}\plus \Varid{f}\;\Varid{x}\plus \Varid{after}{}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{}\mid \Varid{otherwise}\mathrel{=}\Varid{helper}\;(\Varid{before}\plus [\mskip1.5mu \Varid{x}\mskip1.5mu])\;\Varid{after}\;(\Varid{j}\mathbin{+}\mathrm{1}){}\<[E]%
+\\
+\>[4]{}\hsindent{1}{}\<[5]%
+\>[5]{}\Varid{helper}\;\anonymous \;[\mskip1.5mu \mskip1.5mu]\;\anonymous \mathrel{=}\Varid{error}\;\text{\tt \char34 Index~out~of~bounds.\char34}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+  Here \text{\tt modify} takes an update action \text{\tt f}\footnote{\text{\tt f} returns a list rather
+  than a single element to prevent curious readers from suggesting to use a
+  boxed array instead of a list.}, an index \text{\tt i}, and a list \text{\tt xs} and returns a
+  new list with the \text{\tt i}'th element replaced with the result of \text{\tt f}.%
+  This function "unpacks" a list, modifies one element, and "packs" the result
+  into a list. If we do a lot of updates, we end up unpacking and packing the
+  list over and over again -- very time-consuming for long lists. Explicitly
+  working with the unpacked representation is bug-prone. A list zipper
+  simplifies this.
+
+  A zipper consists of a focus (alternatively called a hole) and surrounding
+  context:
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\mathbf{data}\;\Conid{Zipper}\;\Varid{a}\mathrel{=}\Conid{Zipper}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\{\mskip1.5mu \Varid{\char95 hole}\mathbin{::}\Varid{a},\Varid{\char95 cxt}\mathbin{::}\,!(\Conid{Context}\;\Varid{a})\mskip1.5mu\}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\mathbf{deriving}\;(\Conid{Show},\Conid{Eq}){}\<[E]%
+\\
+\>[3]{}\mathbf{data}\;\Conid{Context}\;\Varid{a}\mathrel{=}\Conid{Context}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\mathbf{deriving}\;(\Conid{Show},\Conid{Eq}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+  where the \text{\tt ListContext} keeps track of elements to the left and to the right
+  of the focus. We can now define movements:
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\Varid{left}\mathbin{::}\Conid{Zipper}\;\Varid{a}\to \Conid{Maybe}\;(\Conid{Zipper}\;\Varid{a}){}\<[E]%
+\\
+\>[3]{}\Varid{left}\;(\Conid{Zipper}\;\anonymous \;(\Conid{Context}\;[\mskip1.5mu \mskip1.5mu]\;\anonymous ))\mathrel{=}\Conid{Nothing}{}\<[E]%
+\\
+\>[3]{}\Varid{left}\;(\Conid{Zipper}\;\Varid{hole}\;(\Conid{Context}\;(\Varid{l}\mathbin{:}\Varid{ls})\;\Varid{rs}))\mathrel{=}\Conid{Just}\mathbin{\$}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\Conid{Zipper}\;\Varid{l}\;(\Conid{Context}\;\Varid{ls}\;(\Varid{hole}\mathbin{:}\Varid{rs})){}\<[E]%
+\\[\blanklineskip]%
+\>[3]{}\Varid{right}\mathbin{::}\Conid{Zipper}\;\Varid{a}\to \Conid{Maybe}\;(\Conid{Zipper}\;\Varid{a}){}\<[E]%
+\\
+\>[3]{}\Varid{right}\;(\Conid{Zipper}\;\anonymous \;(\Conid{Context}\;\anonymous \;[\mskip1.5mu \mskip1.5mu]))\mathrel{=}\Conid{Nothing}{}\<[E]%
+\\
+\>[3]{}\Varid{right}\;(\Conid{Zipper}\;\Varid{hole}\;(\Conid{Context}\;\Varid{ls}\;(\Varid{r}\mathbin{:}\Varid{rs})))\mathrel{=}\Conid{Just}\mathbin{\$}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\Conid{Zipper}\;\Varid{r}\;(\Conid{Context}\;(\Varid{hole}\mathbin{:}\Varid{ls})\;\Varid{rs}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+  and functions for entering and leaving the zipper:
 \begin{hscode}\SaveRestoreHook
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\Varid{main}\mathbin{::}\Conid{IO}\;(){}\<[E]%
+\>[3]{}\Varid{lzEnter}\mathbin{::}[\mskip1.5mu \Varid{a}\mskip1.5mu]\to \Conid{Maybe}\;(\Conid{Zipper}\;\Varid{a}){}\<[E]%
 \\
-\>[3]{}\Varid{main}\mathrel{=}\Varid{putStrLn}\;\text{\tt \char34 Hello~world!\char34}{}\<[E]%
+\>[3]{}\Varid{lzEnter}\;[\mskip1.5mu \mskip1.5mu]\mathrel{=}\Conid{Nothing}{}\<[E]%
+\\
+\>[3]{}\Varid{lzEnter}\;(\Varid{x}\mathbin{:}\Varid{xs})\mathrel{=}\Conid{Just}\mathbin{\$}\Conid{Zipper}\;\Varid{x}\;(\Conid{Context}\;[\mskip1.5mu \mskip1.5mu]\;\Varid{xs}){}\<[E]%
+\\[\blanklineskip]%
+\>[3]{}\Varid{lzLeave}\mathbin{::}\Conid{Zipper}\;\Varid{a}\to [\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
+\\
+\>[3]{}\Varid{lzLeave}\;(\Conid{Zipper}\;\Varid{hole}\;(\Conid{Context}\;\Varid{ls}\;\Varid{rs}))\mathrel{=}\Varid{reverse}\;\Varid{ls}\plus \Varid{hole}\mathbin{:}\Varid{rs}{}\<[E]%
 \ColumnHook
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Functional Zippers}
-  The zipper data structure was originally conceived by
-  Huet\cite{huet1997zipper} to solve the problem of representing a tree together
+  Finally, we define a local version of our \text{\tt modify} function (\TODO{Boy, is
+  this function ugly...})
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{19}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\Varid{lzModify}\mathbin{::}(\Varid{a}\to [\mskip1.5mu \Varid{a}\mskip1.5mu])\to \Conid{Zipper}\;\Varid{a}\to \Conid{Maybe}\;(\Conid{Zipper}\;\Varid{a}){}\<[E]%
+\\
+\>[3]{}\Varid{lzModify}\;\Varid{f}\;(\Conid{Zipper}\;\Varid{hole}\;(\Conid{Context}\;\Varid{ls}\;\Varid{rs}))\mathrel{=}\mathbf{case}\;\Varid{f}\;\Varid{hole}\;\mathbf{of}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}(\Varid{x}\mathbin{:}\Varid{xs})\to \Conid{Just}\mathbin{\$}\Conid{Zipper}\;\Varid{x}\;(\Conid{Context}\;\Varid{ls}\;(\Varid{xs}\plus \Varid{rs})){}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}[\mskip1.5mu \mskip1.5mu]{}\<[14]%
+\>[14]{}\to \mathbf{case}\;\Varid{rs}\;\mathbf{of}{}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{}(\Varid{r}\mathbin{:}\Varid{rs'})\to \Conid{Just}\mathbin{\$}\Conid{Zipper}\;\Varid{r}\;(\Conid{Context}\;\Varid{ls}\;\Varid{rs'}){}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{}[\mskip1.5mu \mskip1.5mu]{}\<[17]%
+\>[17]{}\to \mathbf{case}\;\Varid{ls}\;\mathbf{of}{}\<[E]%
+\\
+\>[7]{}\hsindent{2}{}\<[9]%
+\>[9]{}(\Varid{l}\mathbin{:}\Varid{ls'})\to \Conid{Just}\mathbin{\$}\Conid{Zipper}\;\Varid{l}\;(\Conid{Context}\;\Varid{ls'}\;\Varid{rs}){}\<[E]%
+\\
+\>[7]{}\hsindent{2}{}\<[9]%
+\>[9]{}[\mskip1.5mu \mskip1.5mu]{}\<[19]%
+\>[19]{}\to \Conid{Nothing}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+  using which we can perform multiple update efficiently and with minimal code
+  bloat:
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\Varid{modifyExample}\mathbin{::}\Conid{IO}\;(){}\<[E]%
+\\
+\>[3]{}\Varid{modifyExample}\mathrel{=}\Varid{print}\mathbin{\$}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\Varid{lzEnter}\mathord{@}\Conid{Int}\mathbin{>=>}\Varid{right}{}\<[E]%
+\\
+\>[5]{}\hsindent{13}{}\<[18]%
+\>[18]{}\mathbin{>=>}\Varid{right}{}\<[E]%
+\\
+\>[5]{}\hsindent{13}{}\<[18]%
+\>[18]{}\mathbin{>=>}\Varid{lzModify}\;(\Varid{const}\;[\mskip1.5mu \mskip1.5mu]){}\<[E]%
+\\
+\>[5]{}\hsindent{13}{}\<[18]%
+\>[18]{}\mathbin{>=>}\Varid{lzModify}\;(\Varid{return}\mathbin{\circ}(\mathbin{+}\mathrm{1})){}\<[E]%
+\\
+\>[5]{}\hsindent{13}{}\<[18]%
+\>[18]{}\mathbin{>=>}\Varid{left}{}\<[E]%
+\\
+\>[5]{}\hsindent{13}{}\<[18]%
+\>[18]{}\mathbin{>=>}\Varid{lzModify}\;(\Varid{return}\mathbin{\circ}\Varid{negate}){}\<[E]%
+\\
+\>[5]{}\hsindent{13}{}\<[18]%
+\>[18]{}\mathbin{>=>}\Varid{return}\mathbin{\circ}\Varid{lzLeave}\mathbin{\$}{}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{}[\mskip1.5mu \mathrm{1},\mathrm{2},\mathrm{3},\mathrm{4},\mathrm{5}\mskip1.5mu]{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{comment}
+  to solve the problem of representing a tree together
   with a subtree that is the focus of attention, where that focus may move left,
   right, up or down the tree. Bla-bla-bla...
+\end{comment}
 
@@
   Application to lists...
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\mathbf{data}\;\Conid{Path}\;\Varid{a}\mathrel{=}\Conid{Path}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
-\\[\blanklineskip]%
-\>[3]{}\mathbf{data}\;\Conid{Zipper}\;\Varid{a}\mathrel{=}\Conid{Zipper}\,!(\Conid{Path}\;\Varid{a})\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
 

### 2018-09-11 doc/Paper.tex
Update PDF to 61f26b5a7e2e62a847a71aa3461d2659dad37613

--- 2018-08-26 doc/Paper.tex
@@
 \usepackage{amsmath}
+\usepackage{calc}
 \usepackage{color}
@@
 \usepackage{wrapfig}
+\usepackage{pbox}
+\usepackage{tikz}
+\usetikzlibrary{arrows,arrows.meta}
 
@@
 
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
 %
@@
 \newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
 \newcommand{\NB}{\textbf{NB}}
@@
 
+\newcommand{\WithMath}[2]{{\parbox{\widthof{#1}}{\centering$#2$}}}
+
 % A command for declaring todos
 \newcommand{\TODO}[1]{{\color[rgb]{1,0,0}\textbf{TODO:}\textit{#1}}}
+
+\newcommand{\mimplies}{%
+  \hspace{2pt}\makebox[10pt][l]{%
+    \tikz[transform canvas={scale=0.5}]{%
+    \draw [double distance=3pt,
+           line width=0.65pt,
+           arrows={Implies[reversed]}-{Implies}] (0,0.17) -- (0.7,0.17);%
+    }%
+  }%
+  \hspace{2pt}%
+}
 
@@
 
+$\mimplies\Rightarrow\implies$
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Functional Zippers}
+  Zipper is a data structure commonly used in functional programming for
+  traversal with fast local updates. The zipper data structure was originally
+  conceived by Huet\cite{huet1997zipper} in the context of trees. We will,
+  however, first consider a simpler problem: a bidirectional list traversal.
+
+  Suppose that we would like to update a list at a specific position:
 \begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\Varid{main}\mathbin{::}\Conid{IO}\;(){}\<[E]%
+~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]
 \\
-\>[3]{}\Varid{main}\mathrel{=}\Varid{putStrLn}\;\text{\tt \char34 Hello~world!\char34}{}\<[E]%
+\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0
+\\
+\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j
+\\
+\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after
+\\
+\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)
+\\
+\texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
+\ColumnHook
+\end{hscode}\resethooks
+  Here \text{\tt modify} takes an update action \text{\tt f}\footnote{\text{\tt f} returns a list rather
+  than a single element to prevent curious readers from suggesting to use a
+  boxed array instead of a list.}, an index \text{\tt i}, and a list \text{\tt xs} and returns a
+  new list with the \text{\tt i}'th element replaced with the result of \text{\tt f}.%
+  This function "unpacks" a list, modifies one element, and "packs" the result
+  into a list. If we do a lot of updates, we end up unpacking and packing the
+  list over and over again -- very time-consuming for long lists. Explicitly
+  working with the unpacked representation is bug-prone. A list zipper
+  simplifies this.
+
+  A zipper consists of a focus (alternatively called a hole) and surrounding
+  context:
+\begin{hscode}\SaveRestoreHook
+~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
+\\
+\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
+\\
+\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
+\\
+\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
+\ColumnHook
+\end{hscode}\resethooks
+  where the \text{\tt ListContext} keeps track of elements to the left and to the right
+  of the focus. We can now define movements:
+\begin{hscode}\SaveRestoreHook
+~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
+\\
+\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}
+\\
+\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
+\\
+\texfamily 
+\\
+\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
+\\
+\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}
+\\
+\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
+\ColumnHook
+\end{hscode}\resethooks
+  and functions for entering and leaving the zipper:
+\begin{hscode}\SaveRestoreHook
+~~lzEnter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
+\\
+\texfamily ~~lzEnter~[]~~~~~~~=~{\itshape Nothing}
+\\
+\texfamily ~~lzEnter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
+\\
+\texfamily 
+\\
+\texfamily ~~lzLeave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
+\\
+\texfamily ~~lzLeave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
 \ColumnHook
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Functional Zippers}
-  The zipper data structure was originally conceived by
-  Huet\cite{huet1997zipper} to solve the problem of representing a tree together
-  with a subtree that is the focus of attention, where that focus may move left,
-  right, up or down the tree. Bla-bla-bla...
-
-  Application to binary trees...
+  Finally, we define a local version of our \text{\tt modify} function (\TODO{Boy, is
+  this function ugly...})
 \begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\mathbf{data}\;\Conid{Tree}\;\Varid{a}{}\<[E]%
+~~lzModify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
 \\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathrel{=}\Conid{Fork}\;(\Conid{Tree}\;\Varid{a})\;(\Conid{Tree}\;\Varid{a}){}\<[E]%
+\texfamily ~~lzModify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
 \\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mid \Conid{Leaf}\,!\Varid{a}{}\<[E]%
-\\[\blanklineskip]%
-\>[3]{}\mathbf{data}\;\Conid{Path}\;\Varid{a}{}\<[E]%
+\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))
 \\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathrel{=}\Conid{Top}{}\<[E]%
+\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}
 \\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mid \Conid{Left}\,!(\Conid{Path}\;\Varid{a})\;(\Conid{Tree}\;\Varid{a}){}\<[E]%
+\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')
 \\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mid \Conid{TreeRight}\;(\Conid{Tree}\;\Varid{a})\,!(\Conid{Path}\;\Varid{a}){}\<[E]%
-\\[\blanklineskip]%
-\>[3]{}\mathbf{data}\;\Conid{Zipper}\;\Varid{a}\mathrel{=}\Conid{Zipper}\,!(\Conid{Path}\;\Varid{a})\;(\Conid{Tree}\;\Varid{a}){}\<[E]%
+\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}
+\\
+\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)
+\\
+\texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
 \ColumnHook
@@
 
-  Application to lists...
+  \newpage
+  using which we can perform multiple updates efficiently and with minimal code
+  bloat\footnote{%
+  Operator \mimplies comes from \text{\tt Control\char46{}Monad} module in \text{\tt base}
+  and has the following signature:
 \begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\mathbf{data}\;\Conid{Path}\;\Varid{a}\mathrel{=}\Conid{Path}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
-\\[\blanklineskip]%
-\>[3]{}\mathbf{data}\;\Conid{Zipper}\;\Varid{a}\mathrel{=}\Conid{Zipper}\,!(\Conid{Path}\;\Varid{a})\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
+~~(\mimplies)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
+\ColumnHook
+\end{hscode}\resethooks
+  }:
+\begin{hscode}\SaveRestoreHook
+~~modifyExample~::~{\itshape IO}~()
+\\
+\texfamily ~~modifyExample~=~print~\$
+\\
+\texfamily ~~~~lzEnter~@{\itshape Int}~\mimplies~right
+\\
+\texfamily ~~~~~~~~~~~~~~~~~\mimplies~right
+\\
+\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(const~[])
+\\
+\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(return~\WithMath{.}{\circ}~(+1))
+\\
+\texfamily ~~~~~~~~~~~~~~~~~\mimplies~left
+\\
+\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(return~\WithMath{.}{\circ}~negate)
+\\
+\texfamily ~~~~~~~~~~~~~~~~~\mimplies~return~\WithMath{.}{\circ}~lzLeave~\$
+\\
+\texfamily ~~~~~~[1,~2,~3,~4,~5]
 \ColumnHook
@@
 
-  Generic zipper...
+  Consider now a binary tree data structure:
+\begin{hscode}\SaveRestoreHook
+~~{\bfseries data}~{\itshape Tree}~a
+\\
+\texfamily ~~~~=~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
+\\
+\texfamily ~~~~|~{\itshape Leaf}~!a
+\\
+\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
+\ColumnHook
+\end{hscode}\resethooks
+  A binary tree zipper is slightly more insteresting than the list zipper,
+  because we can move up and down the tree as well. The zipper again consists of
+  a hole (a subtree we are focused on) and its surrounding context (a path from
+  the hole to the root of the tree):
+\begin{hscode}\SaveRestoreHook
+~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~({\itshape Tree}~a),~\char95 cxt~::~!({\itshape Context}~a)~\char125 
+\\
+\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
+\\
+\texfamily 
+\\
+\texfamily ~~{\bfseries data}~{\itshape Context}~a
+\\
+\texfamily ~~~~=~{\itshape Top}
+\\
+\texfamily ~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
+\\
+\texfamily ~~~~|~{\itshape TreeRight}~({\itshape Tree}~a)~!({\itshape Context}~a)
+\\
+\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
+\ColumnHook
+\end{hscode}\resethooks
+  To move the zipper down, we "unpack" the current hole:
+\begin{hscode}\SaveRestoreHook
+~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
+\\
+\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~\char95 )~=~{\itshape Nothing}
+\\
+\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape TreeRight}~l~cxt)
+\ColumnHook
+\end{hscode}\resethooks
+  \text{\tt TreeContext} stores everything we need to reconstruct the hole, and \text{\tt tzUp} does
+  exactly that:
+\begin{hscode}\SaveRestoreHook
+~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
+\\
+\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}
+\\
+\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
+\\
+\texfamily ~~up~({\itshape Zipper}~r~({\itshape TreeRight}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
+\ColumnHook
+\end{hscode}\resethooks
+  Implementation of \text{\tt tzLeft}, \text{\tt tzRight}, and \text{\tt tzEnter} is very similar to the
+  List zipper case and is left as an exercise for the reader. \text{\tt tzLeave} differs
+  slightly in that we now move all the way up rather than left:
+\begin{hscode}\SaveRestoreHook
+~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a
+\\
+\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}
+\\
+\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'
+\\
+\texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
+\ColumnHook
+\end{hscode}\resethooks
 
-  An application of generic zipper that we will consider is embedding of
+  The list and binary tree zipper we have considered here are homogeneous
+  zippers: the type of focus does not change upon zipper movement. Such zippers
+  can be a very useful abstraction. For example, a well-known window manager
+  XMonad\cite{xmonad} uses a rose tree zipper to track the window under focus.
+  For other tasks, however, one might need to traverse heterogeneous structures.
+  A zipper that can accomodate such needs is usually called a generic zipper as
+  it relies only on the generic structure of Algebraic Data Types (ADTs). One
+  can view an ADT as an Abstract Syntax Tree (AST) where each node is a Haskell
+  constructor rather than a syntax construct.
+
+  The generic zipper we will use is very similar to the one presented
+  in\cite{adams2010syz}. The most common technique in Haskell for supporting
+  heterogeneous types is Existential Quantification. However, not every type can
+  act as a hole. To support moving down the tree, we need the hole to be
+  \textit{dissectible}, i.e. we would like to be able to dissect the value into
+  the constructor and its arguments. Even though \text{\tt Data\char46{}Data\char46{}gfold} allows us to achieve
+  this, we define out own typeclass which additionally allows us to propagate
+  down arbitrary constraints:
+\begin{hscode}\SaveRestoreHook
+~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}
+\\
+\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a
+\\
+\texfamily 
+\\
+\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}
+\\
+\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b
+\\
+\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
+\ColumnHook
+\end{hscode}\resethooks
+  For example, here is how we can make \text{\tt Tree} an instance of \text{\tt Disssectible}
+\begin{hscode}\SaveRestoreHook
+~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}
+\\
+\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r
+\\
+\texfamily ~~~~dissect~x~=~{\itshape LOne}~x
+\ColumnHook
+\end{hscode}\resethooks
+  We can unpack a \text{\tt Fork} and the zipper will thus be able to go down. \text{\tt Leaf}s,
+  however, are left untouched and trying to go down from a \text{\tt Leaf} will return
+  \text{\tt Nothing}.
+
+  To allow the zipper to move left and right, we need a means to encode
+  arguments to the right of the hole. Following Adams et al, we define a GADT
+  representing constructor arguments to the right of the hole:
+\begin{hscode}\SaveRestoreHook
+~~{\bfseries data}~{\itshape Right}~c~provides~r~{\bfseries where}
+\\
+\texfamily ~~~~{\itshape RNil}~~::~{\itshape Right}~c~r~r
+\\
+\texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~b~\WithMath{->}{\rightarrow}~{\itshape Right}~c~provides~r~\WithMath{->}{\rightarrow}~{\itshape Right}~c~(b~\WithMath{->}{\rightarrow}~provides)~r
+\ColumnHook
+\end{hscode}\resethooks
+  For example, for a tuple \text{\tt \char40{}Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char41{}}, we can have
+\begin{hscode}\SaveRestoreHook
+~~lefts~=~{\itshape LOne}~(,,,,,)~`{\itshape LCons}`~1~`{\itshape LCons}`~2~`{\itshape LCons}`~3
+\\
+\texfamily ~~hole~=~4
+\\
+\texfamily ~~rights~=~5~`{\itshape RCons}`~6~`{\itshape RCons}`~{\itshape RNil}
+\ColumnHook
+\end{hscode}\resethooks
+  generalising this a little, we arrive at:
+\begin{hscode}\SaveRestoreHook
+~~{\bfseries data}~{\itshape LocalContext}~c~hole~rights~parent~=
+\\
+\texfamily ~~~~{\itshape LocalContext}~!({\itshape Left}~c~(hole~\WithMath{->}{\rightarrow}~rights))~!({\itshape Right}~c~rights~parent)
+\\
+\texfamily 
+\\
+\texfamily ~~{\bfseries data}~{\itshape Context}~::~({\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}
+\\
+\texfamily ~~~~{\itshape RootContext}~::~\char'24~c~root.~{\itshape Context}~c~root~root
+\\
+\texfamily ~~~~(:>)~::~\char'24~c~parent~root~hole~rights.~(c~parent,~{\itshape Dissectible}~c~parent)
+\\
+\texfamily ~~~~~~~~~=>~!({\itshape Context}~c~parent~root)
+\\
+\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\rmfamily\enskip\{-\# UNPACK  \#-\}\enskip}~!({\itshape LocalContext}~c~hole~rights~parent)
+\\
+\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Context}~c~hole~root
+\ColumnHook
+\end{hscode}\resethooks
+  And just like before the \text{\tt Zipper} is a product of the hole and context:
+\begin{hscode}\SaveRestoreHook
+~~{\bfseries data}~{\itshape Zipper}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(root~::~{\itshape Type})~=
+\\
+\texfamily ~~~~\char'24~hole.~(c~hole,~{\itshape Dissectible}~c~hole)~=>
+\\
+\texfamily ~~~~~~{\itshape Zipper}~\char123 ~\char95 zHole~::~!hole
+\\
+\texfamily ~~~~~~~~~~~~~,~\char95 zCxt~~::~!({\itshape Context}~c~hole~root)
+\\
+\texfamily ~~~~~~~~~~~~~\char125 
+\ColumnHook
+\end{hscode}\resethooks
+
+  Implementation of movements is quite straightforward and is left out. Please,
+  refer to (\TODO{github repo}) for complete code.
+
+  We now consider a rather interesting application of generic zipper: embedding of
   attribute grammars.
@@
 \section{Attribute Grammars}
-  What attribute grammars are...
+  Attribute grammars (AGs) are an extension of context-free grammars that allow
+  to specify context-sensitive syntax as well as the semantics. AGs achieve it
+  by associating a set of attributes with each grammar symbol. These attributes
+  are defined using evaluation rules assiciated with production rules of the
+  context-free grammar.
 
-  Repmin as two traversals...
+  Attributes are then usually divided into two disjoint sets: synthesized
+  attributes and the inherited attributes. Such distinction is required for the
+  construction of a dependency graph. It is then used for specification of the
+  evaluation order and detection of circularity. In the zipper-based embedding of
+  attribute grammars we make no use of a dependency graph and thus do not divide
+  attributes into classes.
 
-  Repmin as a circular program...
+  We will use repmin\footnote{%
+    The repmin problem: Given a tree of integers, replace every integer with the
+    minimum integer in the tree, in one pass.%
+  } as a running example to illustrate how AGs can be specified using functional
+  zippers. The classical solution is the following circular program:
+\begin{hscode}\SaveRestoreHook
+~~repmin~::~{\itshape Tree}~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Tree}~{\itshape Int}
+\\
+\texfamily ~~repmin~t~=~t'
+\\
+\texfamily ~~~{\bfseries where}
+\\
+\texfamily ~~~~(t',~m')~=~go~t~m'
+\\
+\texfamily ~~~~go~({\itshape Leaf}~x~~~~~~)~m~=~({\itshape Leaf}~m,~x)
+\\
+\texfamily ~~~~go~({\itshape Fork}~xs~ys~~)~m~=~({\itshape Fork}~xs'~ys',~min~m\char95 x~m\char95 y)
+\\
+\texfamily ~~~~~{\bfseries where}
+\\
+\texfamily ~~~~~~(xs',~m\char95 x)~=~go~xs~m
+\\
+\texfamily ~~~~~~(ys',~m\char95 y)~=~go~ys~m
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{comment}
+  Alternatively, we can formulate repmin in terms of an attribute grammar. We
+  use \text{\tt localMin} attribute to represent local minimum of a subtree, \text{\tt globalMin}
+  to represent the global minimum, and \text{\tt updated} for the solution for a subtree.
+\end{comment}
+
+  \TODO{This is a copy-paste} Returning to our example, in
+  \cite{viera2009agsfly}, the authors identified three components for solving
+  repmin: computing the minimal value, passing down the minimal value from the
+  root to the leaves and constructing the resulting tree. In this section, we
+  review the Attribute Grammar for repmin that was introduced by
+  \cite{swierstra1998designing}, and show how each of the three components
+  identified by \cite{viera2009agsfly} in that grammar can be embedded in
+  Haskell using our approach.  The attribute grammar for repmin starts by
+  defining the underlying data structure, i.e., binary leaf trees. The attribute
+  grammar fragments presented in this section follow the standard AG notation of
+  \cite{swierstra1998designing}. In this notation, we straightforwardly use the
+  \text{\tt Tree~Int} datatype from \TODO{where?}.
+
+  Having defined the structure, we need to define functionality. We start by
+  reviewing the AG component that computes the minimal value of a tree:
+\begin{hscode}\SaveRestoreHook
+~~SYN~{\itshape Tree}~{\itshape Int}~[localMin~:~{\itshape Int}]
+\\
+\texfamily ~~SEM~{\itshape Tree}~|~{\itshape Leaf}~lhs.localMin~=~@value
+\\
+\texfamily ~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~@left.localMin~`min`~@right.localMin
+\ColumnHook
+\end{hscode}\resethooks
+  This component declares, using the \text{\tt SYN} keyword, that elements of type \text{\tt Tree}
+  synthesize an attribute \text{\tt localMin} of type \text{\tt Int}. Then, a \text{\tt SEM} sentence
+  defines how \text{\tt localMin} is computed: when the current tree is a leaf, clearly
+  its minimal value is the leaf value itself; when it is the fork of two other
+  trees (the \text{\tt left} and the \text{\tt right} subtrees), we compute the minimal values of
+  each subtree (i.e., their \text{\tt localMin} attribute), and then their minimal value
+  (function \text{\tt min}). In this notation, \text{\tt lhs} refers to the left-hand side symbol
+  of the production and @ prefixes a reference to a field.
+
+  Having implemented the first of the three components that solve repmin, we now
+  consider the remaining two. We start by implementing the construction of the
+  result of repmin, a tree with all leaves being the minimum of the original
+  one.
+\begin{hscode}\SaveRestoreHook
+~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]
+\\
+\texfamily ~~SEM~{\itshape Tree}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin
+\\
+\texfamily ~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated~@right.updated
+\ColumnHook
+\end{hscode}\resethooks
+  The implementation of sres places in each leaf of a tree the value of the ival
+  attribute. This value corresponds to the minimal value of the global tree,
+  that still needs to be passed down to all the nodes in the tree. This
+  corresponds exactly to the third component that we still need to implement. In
+  order to bind the minimal value being computed (attribute smin) with the
+  minimal value that is passed down through the tree (attribute ival), it is
+  common, in the AG setting, to add a new data-type definition,
+
+  In Haskell, these attributes are simply functions:
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{References}
 % As per the LLNCS guidelines

### 2018-09-18 src/Language/AG/Zipper.hs
Zipper initially developed in the memo-AG repo

@@ -1,0 +1,430 @@
+-- {-# LANGUAGE DeriveGeneric #-}
+{-# LANGUAGE DeriveAnyClass #-}
+{-# LANGUAGE NoImplicitPrelude #-}
+{-# LANGUAGE ConstraintKinds #-}
+{-# LANGUAGE FlexibleContexts #-}
+{-# LANGUAGE FlexibleInstances #-}
+{-# LANGUAGE KindSignatures #-}
+{-# LANGUAGE StandaloneDeriving #-}
+{-# LANGUAGE FunctionalDependencies #-}
+-- {-# LANGUAGE InstanceSigs #-}
+{-# LANGUAGE AllowAmbiguousTypes #-}
+-- {-# LANGUAGE UndecidableSuperClasses #-}
+-- {-# LANGUAGE NoMonomorphismRestriction #-}
+-- {-# LANGUAGE UndecidableInstances #-}
+{-# LANGUAGE ScopedTypeVariables #-}
+-- {-# LANGUAGE TypeFamilyDependencies #-}
+{-# LANGUAGE LambdaCase #-}
+{-# LANGUAGE TypeApplications #-}
+{-# LANGUAGE DataKinds #-}
+{-# LANGUAGE RankNTypes #-}
+{-# LANGUAGE GADTs #-}
+{-# LANGUAGE TypeInType #-}
+{-# LANGUAGE TemplateHaskell #-}
+{-# LANGUAGE TypeFamilies #-}
+{-# LANGUAGE TypeOperators #-}
+{-# LANGUAGE ViewPatterns #-}
+{-# LANGUAGE BangPatterns #-}
+{-# LANGUAGE MultiParamTypeClasses #-}
+-- {-# LANGUAGE LiberalTypeSynonyms #-}
+{-# LANGUAGE ExistentialQuantification #-}
+-- {-# LANGUAGE UnicodeSyntax #-}
+{-# LANGUAGE OverloadedLabels #-}
+{-# LANGUAGE PatternSynonyms #-}
+
+-- |
+-- Module      : Zipper
+-- Description : A generic zipper for embedding attribute grammars.
+-- Copyright   : (c) Tom Westerhout, 2018
+-- License     : GPL-3
+-- Maintainer  : t.westerhout@student.ru.nl
+-- Stability   : experimental
+module Language.AG.Zipper
+  ( -- ** Motivation
+    -- $motivation
+
+    -- ** The Zipper
+    Zipper(..)
[…]

### 2018-09-18 doc/Paper.tex
Update PDF to a39a0132dc040c4486ad7794acafbe8b27e6ae2c

@@
 \usepackage{calc}
+\usepackage{csquotes}
 \usepackage{color}
@@
 \usepackage{pbox}
-\usepackage{tikz}
-\usetikzlibrary{arrows,arrows.meta}
 
@@
 
-\newcommand{\WithMath}[2]{{\parbox{\widthof{#1}}{\centering$#2$}}}
+\newcommand{\WithMath}[2]{{\parbox[][][b]{\widthof{#1}}{\centering$#2$}}}
+
+% For the circular variant of repmin
+
+% For the specification of AGs
 
@@
 \newcommand{\TODO}[1]{{\color[rgb]{1,0,0}\textbf{TODO:}\textit{#1}}}
-
-\newcommand{\mimplies}{%
-  \hspace{2pt}\makebox[10pt][l]{%
-    \tikz[transform canvas={scale=0.5}]{%
-    \draw [double distance=3pt,
-           line width=0.65pt,
-           arrows={Implies[reversed]}-{Implies}] (0,0.17) -- (0.7,0.17);%
-    }%
-  }%
-  \hspace{2pt}%
-}
 
@@
 \section{Introduction}\label{sec:introduction}
-
-$\mimplies\Rightarrow\implies$
 
@@
   bloat\footnote{%
-  Operator \mimplies comes from \text{\tt Control\char46{}Monad} module in \text{\tt base}
+  Operator \text{\tt \char62{}\char61{}\char62{}} comes from \text{\tt Control\char46{}Monad} module in \text{\tt base}
   and has the following signature:
 \begin{hscode}\SaveRestoreHook
-~~(\mimplies)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
+~~(>=>)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
 \ColumnHook
@@
 \\
-\texfamily ~~~~lzEnter~@{\itshape Int}~\mimplies~right
+\texfamily ~~~~lzEnter~@{\itshape Int}~>=>~right
 \\
-\texfamily ~~~~~~~~~~~~~~~~~\mimplies~right
+\texfamily ~~~~~~~~~~~~~~~~~>=>~right
 \\
-\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(const~[])
+\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(const~[])
 \\
-\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(return~\WithMath{.}{\circ}~(+1))
+\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~(+1))
 \\
-\texfamily ~~~~~~~~~~~~~~~~~\mimplies~left
+\texfamily ~~~~~~~~~~~~~~~~~>=>~left
 \\
-\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(return~\WithMath{.}{\circ}~negate)
+\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~negate)
 \\
-\texfamily ~~~~~~~~~~~~~~~~~\mimplies~return~\WithMath{.}{\circ}~lzLeave~\$
+\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~lzLeave~\$
 \\
@@
 
-  We will use repmin\footnote{%
-    The repmin problem: Given a tree of integers, replace every integer with the
-    minimum integer in the tree, in one pass.%
-  } as a running example to illustrate how AGs can be specified using functional
-  zippers. The classical solution is the following circular program:
+  Let us consider the repmin\footnote{%
+    The repmin problem:
+    \begin{displayquote}
+      Given a tree of integers, replace every integer with the
+      minimum integer in the tree, in one pass.
+    \end{displayquote}
+  } problem as an example of a problem that requires multiple traversals. The
+  classical solution is the following circular program:
 \begin{hscode}\SaveRestoreHook
@@
 \\
-\texfamily ~~~{\bfseries where}
+\texfamily ~~~~{\bfseries where}~(t',~m')~=~go~t~m'
 \\
-\texfamily ~~~~(t',~m')~=~go~t~m'
+\texfamily ~~~~~~~~~~go~({\itshape Leaf}~x~~~~)~m~=~({\itshape Leaf}~m,~x)
 \\
-\texfamily ~~~~go~({\itshape Leaf}~x~~~~~~)~m~=~({\itshape Leaf}~m,~x)
+\texfamily ~~~~~~~~~~go~({\itshape Fork}~xs~ys)~m~=~({\itshape Fork}~xs'~ys',~min~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$}~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})
 \\
-\texfamily ~~~~go~({\itshape Fork}~xs~ys~~)~m~=~({\itshape Fork}~xs'~ys',~min~m\char95 x~m\char95 y)
+\texfamily ~~~~~~~~~~~~{\bfseries where}~(xs',~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$})~=~go~xs~m
 \\
-\texfamily ~~~~~{\bfseries where}
-\\
-\texfamily ~~~~~~(xs',~m\char95 x)~=~go~xs~m
-\\
-\texfamily ~~~~~~(ys',~m\char95 y)~=~go~ys~m
+\texfamily ~~~~~~~~~~~~~~~~~~(ys',~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})~=~go~ys~m
 \ColumnHook
@@
 
-\begin{comment}
-  Alternatively, we can formulate repmin in terms of an attribute grammar. We
-  use \text{\tt localMin} attribute to represent local minimum of a subtree, \text{\tt globalMin}
-  to represent the global minimum, and \text{\tt updated} for the solution for a subtree.
-\end{comment}
+  Although quite elegant, the code lacks modularity and is very difficult to
+  reason about. Attribute Grammars provide a more modular approach. Viera et
+  al\cite{viera2009agsfly} identified three steps for solving repmin: computing
+  the minimal value, passing it down from the root to the leaves, and
+  constructing the resulting tree. We can associate each step with an
+  attribute\cite{swierstra1998designing}:
+  \begin{itemize}
+  \item A synthesized attribute \text{\texfamily localMin\Sp ::\Sp {\itshape Int}} represents the minimum value
+        of a subtree. Computing the minimal value thus corresponds to evaluation
+        of the \text{\texfamily localMin} attribute for the root tree.
+  \item An inherited attribute \text{\texfamily globalMin\Sp ::\Sp {\itshape Int}} is used to pass down the
+        minimal value.
+  \item Finally, a synthesized attribute \text{\texfamily updated\Sp ::\Sp {\itshape Tree}\Sp {\itshape Int}} is the subtree
+        with leaf values replaced by values of their \text{\texfamily globalMin} attributes. The
+        solution is thus the value of \text{\texfamily updated} attribute for the root tree.
+  \end{itemize}
+  The obtained AG is presented in figure~\ref{fig:repmin-AG}. \TODO{Do we
+  actually need to explain the algorithms here? It seems a little bit childish
+  to explain how to compute the minimum of a binary tree... If we absolutely
+  have to explain stuff, maybe just put it in the caption.}
 
-  \TODO{This is a copy-paste} Returning to our example, in
-  \cite{viera2009agsfly}, the authors identified three components for solving
-  repmin: computing the minimal value, passing down the minimal value from the
-  root to the leaves and constructing the resulting tree. In this section, we
-  review the Attribute Grammar for repmin that was introduced by
-  \cite{swierstra1998designing}, and show how each of the three components
-  identified by \cite{viera2009agsfly} in that grammar can be embedded in
-  Haskell using our approach.  The attribute grammar for repmin starts by
-  defining the underlying data structure, i.e., binary leaf trees. The attribute
-  grammar fragments presented in this section follow the standard AG notation of
-  \cite{swierstra1998designing}. In this notation, we straightforwardly use the
-  \text{\tt Tree~Int} datatype from \TODO{where?}.
-
-  Having defined the structure, we need to define functionality. We start by
-  reviewing the AG component that computes the minimal value of a tree:
+\begin{figure}
+\centering
+\fbox{\parbox{\linewidth}{%
 \begin{hscode}\SaveRestoreHook
@@
 \\
-\texfamily ~~SEM~{\itshape Tree}~|~{\itshape Leaf}~lhs.localMin~=~@value
+\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.localMin~=~@value
 \\
-\texfamily ~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~@left.localMin~`min`~@right.localMin
+\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~min~@left.localMin
+\\
+\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.localMin
+\\
+\texfamily 
+\\
+\texfamily ~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]
+\\
+\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin
+\\
+\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated
+\\
+\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.updated
+\\
+\texfamily 
+\\
+\texfamily 
+\\
+\texfamily ~~INH~{\itshape Tree}~{\itshape Int}~[~globalMin~:~{\itshape Int}~]
+\\
+\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Fork}~left.globalMin~~=~@lhs.globalMin
+\\
+\texfamily ~~~~~~~~~~~~~~~~~~~~~~right.globalMin~=~@lhs.globalMin
+\\
+\texfamily 
+\\
+\texfamily ~~DATA~{\itshape Root}~|~{\itshape Root}~tree~:~{\itshape Tree}~{\itshape Int}
+\\
+\texfamily ~~SEM~~{\itshape Root}~|~{\itshape Root}~tree.globalMin~=~@tree.localMin
 \ColumnHook
 \end{hscode}\resethooks
-  This component declares, using the \text{\tt SYN} keyword, that elements of type \text{\tt Tree}
-  synthesize an attribute \text{\tt localMin} of type \text{\tt Int}. Then, a \text{\tt SEM} sentence
-  defines how \text{\tt localMin} is computed: when the current tree is a leaf, clearly
-  its minimal value is the leaf value itself; when it is the fork of two other
-  trees (the \text{\tt left} and the \text{\tt right} subtrees), we compute the minimal values of
-  each subtree (i.e., their \text{\tt localMin} attribute), and then their minimal value
-  (function \text{\tt min}). In this notation, \text{\tt lhs} refers to the left-hand side symbol
-  of the production and @ prefixes a reference to a field.
+%
+}}
+\caption{%
+  Attribute grammar for repmin. The syntax is closely mirrors the one used
+  in\cite{swierstra1998designing}. \text{\texfamily SYN} and \text{\texfamily INH} introduce synthesized
+  and inherited attributes respectively. \text{\texfamily SEM} is used for defining semantic
+  rules. A new data type \text{\texfamily {\itshape Root}} is introduced as it is common in the AG setting
+  to ``connect'' \text{\texfamily localMin} with \text{\texfamily globalMin}.
+}
+\label{fig:repmin-AG}
+\end{figure}
 
-  Having implemented the first of the three components that solve repmin, we now
-  consider the remaining two. We start by implementing the construction of the
-  result of repmin, a tree with all leaves being the minimum of the original
-  one.
+  We now move on to embed this attribute grammar into Haskell. Semantic rules
+  simply become functions, which, given a zipper, return values of the
+  attributes. For example,
 \begin{hscode}\SaveRestoreHook
-~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]
+~~localMin~::~{\itshape {\itshape Lib}.Zipper}~{\itshape Cxt}~{\itshape Attributes}~({\itshape Tree}~{\itshape Int})~\WithMath{->}{\rightarrow}~{\itshape Int}
 \\
-\texfamily ~~SEM~{\itshape Tree}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin
+\texfamily ~~localMin~z@({\itshape {\itshape Lib}.Zipper}~hole)~=~{\bfseries case}~whereami~hole~{\bfseries of}
 \\
-\texfamily ~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated~@right.updated
+\texfamily ~~~~{\itshape C\char95 Leaf}~\WithMath{->}{\rightarrow}~{\bfseries let}~({\itshape Leaf}~x)~=~hole~{\bfseries in}~x
+\\
+\texfamily ~~~~{\itshape C\char95 Fork}~\WithMath{->}{\rightarrow}~min~(localMin~\$~{\itshape Lib}.unsafeChild~0~z)
+\\
+\texfamily ~~~~~~~~~~~~~~~~~~(localMin~\$~{\itshape Lib}.unsafeChild~1~z)
 \ColumnHook
 \end{hscode}\resethooks
-  The implementation of sres places in each leaf of a tree the value of the ival
-  attribute. This value corresponds to the minimal value of the global tree,
-  that still needs to be passed down to all the nodes in the tree. This
-  corresponds exactly to the third component that we still need to implement. In
-  order to bind the minimal value being computed (attribute smin) with the
-  minimal value that is passed down through the tree (attribute ival), it is
-  common, in the AG setting, to add a new data-type definition,
 
   In Haskell, these attributes are simply functions:
-
-  Repmin as an AG...
 

### 2018-09-18 src/Language/AG/Constraint.hs
(&&&) for Contraints

@@ -1,0 +1,34 @@
+{-# LANGUAGE ConstraintKinds #-}
+{-# LANGUAGE ExplicitNamespaces #-}
+{-# LANGUAGE MultiParamTypeClasses #-}
+{-# LANGUAGE UndecidableSuperClasses #-}
+{-# LANGUAGE TypeInType #-}
+{-# LANGUAGE TypeOperators #-}
+{-# OPTIONS_HADDOCK show-extensions #-}
+
+-- |
+-- Module      : Language.AG.Contraint
+-- Description : Utilities related to 'Contraint's
+-- Copyright   : (c) Tom Westerhout, 2018
+-- License     : BSD-3
+-- Maintainer  : t.westerhout@student.ru.nl
+-- Stability   : experimental
+-- Portability : GHC
+module Language.AG.Constraint
+  ( -- A utility typeclass that acts as a logical *and* for types of kind @'Type' -> 'Contraint'@.
+    --
+    -- @
+    -- type OrdShow = Show && Ord
+    --
+    -- printMin :: OrdShow a => a -> a -> IO ()
+    -- printMin x y = if x <= y then print x else print y
+    -- @
+    type (&&&)
+  ) where
+
+import Data.Kind
+
+class (a c, b c) => (&&&) (a :: Type -> Constraint) (b :: Type -> Constraint) c
+infixl 4 &&&
+
+instance (a c, b c) => (a &&& b) c

### 2018-09-18 src/Language/AG/Zipper.hs
Fix a type in the license

@@ -36,7 +36,7 @@
 -- Module      : Zipper
 -- Description : A generic zipper for embedding attribute grammars.
 -- Copyright   : (c) Tom Westerhout, 2018
--- License     : GPL-3
+-- License     : BSD-3
 -- Maintainer  : t.westerhout@student.ru.nl
 -- Stability   : experimental
 module Language.AG.Zipper

### 2018-09-18 doc/Paper.tex
Update PDF to daf8d2c4d984da3ed6239fe13805c50e32749742

@@
 % For the specification of AGs
+
+% Tree positions
 
@@
 \begin{hscode}\SaveRestoreHook
-~~localMin~::~{\itshape {\itshape Lib}.Zipper}~{\itshape Cxt}~{\itshape Attributes}~({\itshape Tree}~{\itshape Int})~\WithMath{->}{\rightarrow}~{\itshape Int}
+~~localMin~::~{\itshape Zipper}~({\itshape WhereAmI}~{\itshape Position})~({\itshape Tree}~{\itshape Int})~\WithMath{->}{\rightarrow}~{\itshape Int}
 \\
-\texfamily ~~localMin~z@({\itshape {\itshape Lib}.Zipper}~hole)~=~{\bfseries case}~whereami~hole~{\bfseries of}
+\texfamily ~~localMin~z@({\itshape Zipper}~hole~\char95 )~=~{\bfseries case}~whereami~hole~{\bfseries of}
 \\
-\texfamily ~~~~{\itshape C\char95 Leaf}~\WithMath{->}{\rightarrow}~{\bfseries let}~({\itshape Leaf}~x)~=~hole~{\bfseries in}~x
+\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Leaf}~x~=~hole~{\bfseries in}~x
 \\
-\texfamily ~~~~{\itshape C\char95 Fork}~\WithMath{->}{\rightarrow}~min~(localMin~\$~{\itshape Lib}.unsafeChild~0~z)
+\texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Just}~l~=~child~0~z;~{\itshape Just}~r~=~child~1~z
 \\
-\texfamily ~~~~~~~~~~~~~~~~~~(localMin~\$~{\itshape Lib}.unsafeChild~1~z)
+\texfamily ~~~~~~~~~~~~~~{\bfseries in}~min~(localMin~l)~(localMin~r)
 \ColumnHook
 \end{hscode}\resethooks
+
+  Apart from the type signature, the code is pretty straightforward and closely
+  mirrors the AG we defined earlier. \text{\texfamily whereami} function allows us to ``look
+  around'' and returns the position of the zipper. Thus the \text{\texfamily {\bfseries case}} corresponds
+  to the pattern matches on the left of the vertical bars on
+  figure~\ref{fig:repmin-AG}.
+
+  Position of the zipper is encoded using the following GADT which can be
+  generated automatically using Template Haskell:
+\begin{hscode}\SaveRestoreHook
+~~{\bfseries data}~{\itshape Position}~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}
+\\
+\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})
+\\
+\texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})
+\ColumnHook
+\end{hscode}\resethooks
+  Parametrization on the type of the hole allows the code like
+  \text{\texfamily {\bfseries let}\Sp {\itshape Leaf}\Sp x\Sp =\Sp hole\Sp {\bfseries in}\Sp x} to typecheck even though the generic zipper itself knows close
+  to nothing about the type of the hole.
 

### 2018-09-18 doc/Paper.tex
Update PDF to 43d2717eec03c9db275147b4e66fb12fc705110f

@@
 %
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
 %
@@
   Suppose that we would like to update a list at a specific position:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]
@@
 \texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
-\ColumnHook
-\end{hscode}\resethooks
-  Here \text{\tt modify} takes an update action \text{\tt f}\footnote{\text{\tt f} returns a list rather
+\end{tabbing}
+  Here \text{\ttfamily modify} takes an update action \text{\ttfamily f}\footnote{\text{\ttfamily f} returns a list rather
   than a single element to prevent curious readers from suggesting to use a
-  boxed array instead of a list.}, an index \text{\tt i}, and a list \text{\tt xs} and returns a
-  new list with the \text{\tt i}'th element replaced with the result of \text{\tt f}.%
+  boxed array instead of a list.}, an index \text{\ttfamily i}, and a list \text{\ttfamily xs} and returns a
+  new list with the \text{\ttfamily i}'th element replaced with the result of \text{\ttfamily f}.%
   This function "unpacks" a list, modifies one element, and "packs" the result
@@
   context:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
@@
 \texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
-\ColumnHook
-\end{hscode}\resethooks
-  where the \text{\tt ListContext} keeps track of elements to the left and to the right
+\end{tabbing}
+  where the \text{\ttfamily ListContext} keeps track of elements to the left and to the right
   of the focus. We can now define movements:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
@@
 \texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
   and functions for entering and leaving the zipper:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~lzEnter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
@@
 \texfamily ~~lzLeave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
 
-  Finally, we define a local version of our \text{\tt modify} function (\TODO{Boy, is
+  Finally, we define a local version of our \text{\ttfamily modify} function (\TODO{Boy, is
   this function ugly...})
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~lzModify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
@@
 \texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
 
@@
   bloat\footnote{%
-  Operator \text{\tt \char62{}\char61{}\char62{}} comes from \text{\tt Control\char46{}Monad} module in \text{\tt base}
+  Operator \text{\ttfamily \char62{}\char61{}\char62{}} comes from \text{\ttfamily Control\char46{}Monad} module in \text{\ttfamily base}
   and has the following signature:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~(>=>)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
   }:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~modifyExample~::~{\itshape IO}~()
@@
 \texfamily ~~~~~~[1,~2,~3,~4,~5]
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
 
   Consider now a binary tree data structure:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~{\bfseries data}~{\itshape Tree}~a
@@
 \texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
   A binary tree zipper is slightly more insteresting than the list zipper,
@@
   the hole to the root of the tree):
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~({\itshape Tree}~a),~\char95 cxt~::~!({\itshape Context}~a)~\char125 
@@
 \texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
   To move the zipper down, we "unpack" the current hole:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
@@
 \texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape TreeRight}~l~cxt)
-\ColumnHook
-\end{hscode}\resethooks
-  \text{\tt TreeContext} stores everything we need to reconstruct the hole, and \text{\tt tzUp} does
+\end{tabbing}
+  \text{\ttfamily TreeContext} stores everything we need to reconstruct the hole, and \text{\ttfamily tzUp} does
   exactly that:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
@@
 \texfamily ~~up~({\itshape Zipper}~r~({\itshape TreeRight}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
-\ColumnHook
-\end{hscode}\resethooks
-  Implementation of \text{\tt tzLeft}, \text{\tt tzRight}, and \text{\tt tzEnter} is very similar to the
-  List zipper case and is left as an exercise for the reader. \text{\tt tzLeave} differs
+\end{tabbing}
+  Implementation of \text{\ttfamily tzLeft}, \text{\ttfamily tzRight}, and \text{\ttfamily tzEnter} is very similar to the
+  List zipper case and is left as an exercise for the reader. \text{\ttfamily tzLeave} differs
   slightly in that we now move all the way up rather than left:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a
@@
 \texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
 
@@
   \textit{dissectible}, i.e. we would like to be able to dissect the value into
-  the constructor and its arguments. Even though \text{\tt Data\char46{}Data\char46{}gfold} allows us to achieve
+  the constructor and its arguments. Even though \text{\ttfamily Data\char46{}Data\char46{}gfold} allows us to achieve
   this, we define out own typeclass which additionally allows us to propagate
   down arbitrary constraints:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}
@@
 \texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
-\ColumnHook
-\end{hscode}\resethooks
-  For example, here is how we can make \text{\tt Tree} an instance of \text{\tt Disssectible}
-\begin{hscode}\SaveRestoreHook
+\end{tabbing}
+  For example, here is how we can make \text{\ttfamily Tree} an instance of \text{\ttfamily Disssectible}
+\begin{tabbing}\texfamily
 ~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}
@@
 \texfamily ~~~~dissect~x~=~{\itshape LOne}~x
-\ColumnHook
-\end{hscode}\resethooks
-  We can unpack a \text{\tt Fork} and the zipper will thus be able to go down. \text{\tt Leaf}s,
-  however, are left untouched and trying to go down from a \text{\tt Leaf} will return
-  \text{\tt Nothing}.
+\end{tabbing}
+  We can unpack a \text{\ttfamily Fork} and the zipper will thus be able to go down. \text{\ttfamily Leaf}s,
+  however, are left untouched and trying to go down from a \text{\ttfamily Leaf} will return
+  \text{\ttfamily Nothing}.
 
@@
   representing constructor arguments to the right of the hole:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~{\bfseries data}~{\itshape Right}~c~provides~r~{\bfseries where}
@@
 \texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~b~\WithMath{->}{\rightarrow}~{\itshape Right}~c~provides~r~\WithMath{->}{\rightarrow}~{\itshape Right}~c~(b~\WithMath{->}{\rightarrow}~provides)~r
-\ColumnHook
-\end{hscode}\resethooks
-  For example, for a tuple \text{\tt \char40{}Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char41{}}, we can have
-\begin{hscode}\SaveRestoreHook
+\end{tabbing}
+  For example, for a tuple \text{\ttfamily \char40{}Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char41{}}, we can have
+\begin{tabbing}\texfamily
 ~~lefts~=~{\itshape LOne}~(,,,,,)~`{\itshape LCons}`~1~`{\itshape LCons}`~2~`{\itshape LCons}`~3
@@
 \texfamily ~~rights~=~5~`{\itshape RCons}`~6~`{\itshape RCons}`~{\itshape RNil}
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
   generalising this a little, we arrive at:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~{\bfseries data}~{\itshape LocalContext}~c~hole~rights~parent~=
@@
 \texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Context}~c~hole~root
-\ColumnHook
-\end{hscode}\resethooks
-  And just like before the \text{\tt Zipper} is a product of the hole and context:
-\begin{hscode}\SaveRestoreHook
+\end{tabbing}
+  And just like before the \text{\ttfamily Zipper} is a product of the hole and context:
+\begin{tabbing}\texfamily
 ~~{\bfseries data}~{\itshape Zipper}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(root~::~{\itshape Type})~=
@@
 \texfamily ~~~~~~~~~~~~~\char125 
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
 
@@
   classical solution is the following circular program:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~repmin~::~{\itshape Tree}~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Tree}~{\itshape Int}
@@
 \texfamily ~~~~~~~~~~~~~~~~~~(ys',~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})~=~go~ys~m
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
 
@@
 \fbox{\parbox{\linewidth}{%
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~SYN~{\itshape Tree}~{\itshape Int}~[localMin~:~{\itshape Int}]
@@
 \texfamily ~~SEM~~{\itshape Root}~|~{\itshape Root}~tree.globalMin~=~@tree.localMin
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
 %
@@
   attributes. For example,
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~localMin~::~{\itshape Zipper}~({\itshape WhereAmI}~{\itshape Position})~({\itshape Tree}~{\itshape Int})~\WithMath{->}{\rightarrow}~{\itshape Int}
@@
 \texfamily ~~~~~~~~~~~~~~{\bfseries in}~min~(localMin~l)~(localMin~r)
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
 
@@
   generated automatically using Template Haskell:
-\begin{hscode}\SaveRestoreHook
+\begin{tabbing}\texfamily
 ~~{\bfseries data}~{\itshape Position}~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}
@@
 \texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})
-\ColumnHook
-\end{hscode}\resethooks
+\end{tabbing}
   Parametrization on the type of the hole allows the code like
-  \text{\texfamily {\bfseries let}\Sp {\itshape Leaf}\Sp x\Sp =\Sp hole\Sp {\bfseries in}\Sp x} to typecheck even though the generic zipper itself knows close
-  to nothing about the type of the hole.
+  \text{\texfamily {\bfseries let}\Sp {\itshape Leaf}\Sp x\Sp =\Sp hole\Sp {\bfseries in}\Sp x} to typecheck, even though the generic zipper itself knows close
+  to nothing about the type of the hole. It might seem trivial at first, because
+  the binary tree zipper is in fact homogeneous. The ``position trick'' however
+  extends also to heterogeneous zippers which we will encounter in more advanced
+  examples.
 
-  In Haskell, these attributes are simply functions:
+\begin{tabbing}\texfamily
+~~child~::~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Zipper}~cxt~root~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~cxt~root)
+\end{tabbing}
+  \text{\texfamily child\Sp n} moves the zipper to the \text{\texfamily n}'th child, if there is one.
 

### 2018-11-20 doc/Paper.tex
Update PDF to 4b4c412b5d229a27577dcdad334bdfa8c74a5e3b

@@
 
+\newcommand{\ttsub}[2]{{\text{#1}_\text{#2}}}
+
 \begin{document}
@@
 
-  Suppose that we would like to update a list at a specific position:
+  \paragraph{Lists} Suppose that we would like to update a list at a specific
+  position:
 \begin{tabbing}\texfamily
@@
 \end{tabbing}
-  Here \text{\ttfamily modify} takes an update action \text{\ttfamily f}\footnote{\text{\ttfamily f} returns a list rather
+  Here \text{\texfamily modify} takes an update action \text{\texfamily f}\footnote{\text{\texfamily f} returns a list rather
   than a single element to prevent curious readers from suggesting to use a
-  boxed array instead of a list.}, an index \text{\ttfamily i}, and a list \text{\ttfamily xs} and returns a
-  new list with the \text{\ttfamily i}'th element replaced with the result of \text{\ttfamily f}.%
-  This function "unpacks" a list, modifies one element, and "packs" the result
-  into a list. If we do a lot of updates, we end up unpacking and packing the
-  list over and over again -- very time-consuming for long lists. Explicitly
-  working with the unpacked representation is bug-prone. A list zipper
-  simplifies this.
+  boxed array instead of a list.}, an index \text{\texfamily i}, and a list \text{\texfamily xs} and returns a
+  new list with the \text{\texfamily i}'th element replaced with the result of \text{\texfamily f}.%
+  First, we ``unpack'' the list into $\text{\texfamily before\Sp =\Sp [}\ttsub{\text{\texfamily xs}}{\text{\texfamily 0}},\dots
+  \ttsub{\text{\texfamily xs}}{\text{\texfamily i-1}}\text{\texfamily ]}$, $\text{\texfamily x\Sp =\Sp }\ttsub{\text{\texfamily xs}}{\text{\texfamily i}}$, and
+  $\text{\texfamily after\Sp =\Sp [}\ttsub{\text{\texfamily xs}}{\text{\texfamily i+1}},\dots\text{\texfamily ]}$. Then replace \text{\texfamily x} by \text{\texfamily f(x)} and
+  finally ``pack'' the result back into a single list. If we do a lot of
+  updates, we end up unpacking and packing the list over and over again -- very
+  time-consuming for long lists. We would thus like to benefit from fusion
+  without explicitly working with the unpacked representation as it is
+  bug-prone. A list zipper provides a way to achieve this.
 
   A zipper consists of a focus (alternatively called a hole) and surrounding
-  context:
+  context\footnote{`\text{\texfamily !}' in front of the type of \text{\texfamily \char95 cxt} is a
+  strictness annotation enabled by the \text{\ttfamily BangPatterns} extension. Context can be
+  seen as a path from the to the current focus and is always finite, i.e. there
+  is no reason for it to be lazy.}:
 \begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
-\\
-\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
+~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
 \\
 \texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
-\\
-\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
 \end{tabbing}
-  where the \text{\ttfamily ListContext} keeps track of elements to the left and to the right
-  of the focus. We can now define movements:
+  where the \text{\texfamily {\itshape Context}} keeps track of elements to the left (\text{\texfamily before} in
+  \text{\texfamily modify}) and to the right (\text{\texfamily after} in \text{\texfamily modify}) of the focus. We can now
+  define movements:
 \begin{tabbing}\texfamily
@@
 \\
-\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
+\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$
+\\
+\texfamily ~~~~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
 \\
@@
 \\
-\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
+\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$
+\\
+\texfamily ~~~~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
 \end{tabbing}
@@
 \begin{tabbing}\texfamily
-~~lzEnter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
+~~enter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
 \\
-\texfamily ~~lzEnter~[]~~~~~~~=~{\itshape Nothing}
+\texfamily ~~enter~[]~~~~~~~=~{\itshape Nothing}
 \\
-\texfamily ~~lzEnter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
+\texfamily ~~enter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
 \\
@@
 \\
-\texfamily ~~lzLeave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
+\texfamily ~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
 \\
-\texfamily ~~lzLeave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
+\texfamily ~~leave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
 \end{tabbing}
 
-  Finally, we define a local version of our \text{\ttfamily modify} function (\TODO{Boy, is
+  Finally, we define a local version of our \text{\texfamily modify} function (\TODO{Boy, is
   this function ugly...})
 \begin{tabbing}\texfamily
-~~lzModify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
+~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
 \\
-\texfamily ~~lzModify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
+\texfamily ~~modify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
 \\
@@
 \end{tabbing}
-
-  \newpage
   using which we can perform multiple updates efficiently and with minimal code
   bloat\footnote{%
-  Operator \text{\ttfamily \char62{}\char61{}\char62{}} comes from \text{\ttfamily Control\char46{}Monad} module in \text{\ttfamily base}
-  and has the following signature:
+  Operator \text{\ttfamily \char62{}\char61{}\char62{}} comes from \text{\texfamily {\itshape {\itshape Control}.Monad}} module in \text{\ttfamily base}
+  and has the following signature:%
 \begin{tabbing}\texfamily
@@
 \end{tabbing}
+%
   }:
@@
 \\
-\texfamily ~~~~lzEnter~@{\itshape Int}~>=>~right
+\texfamily ~~~~enter~@{\itshape Int}~>=>~right
 \\
@@
 \\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(const~[])
+\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(const~[])
 \\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~(+1))
+\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~(+1))
 \\
@@
 \\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~negate)
+\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~negate)
 \\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~lzLeave~\$
-\\
-\texfamily ~~~~~~[1,~2,~3,~4,~5]
+\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~leave~\$~[1,~2,~3,~4,~5]
 \end{tabbing}
 
-  Consider now a binary tree data structure:
+  \paragraph{Trees} Consider now a binary tree data structure:
 \begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Tree}~a
-\\
-\texfamily ~~~~=~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
-\\
-\texfamily ~~~~|~{\itshape Leaf}~!a
-\\
-\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
+~~{\bfseries data}~{\itshape Tree}~a~=~{\itshape Leaf}~!a~|~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
 \end{tabbing}
   A binary tree zipper is slightly more insteresting than the list zipper,
-  because we can move up and down the tree as well. The zipper again consists of
-  a hole (a subtree we are focused on) and its surrounding context (a path from
-  the hole to the root of the tree):
+  because we can move up and down the tree as well as left and right. The zipper
+  again consists of a hole (a subtree we are focused on) and its surrounding
+  context (a path from the hole to the root of the tree):
 \begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~({\itshape Tree}~a),~\char95 cxt~::~!({\itshape Context}~a)~\char125 
+~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~{\itshape Tree}~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
 \\
-\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
+\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Top}
 \\
-\texfamily 
+\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
 \\
-\texfamily ~~{\bfseries data}~{\itshape Context}~a
-\\
-\texfamily ~~~~=~{\itshape Top}
-\\
-\texfamily ~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
-\\
-\texfamily ~~~~|~{\itshape TreeRight}~({\itshape Tree}~a)~!({\itshape Context}~a)
-\\
-\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
+\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Right}~({\itshape Tree}~a)~!({\itshape Context}~a)
 \end{tabbing}
-  To move the zipper down, we "unpack" the current hole:
+  To move the zipper down, we ``unpack'' the current hole:
 \begin{tabbing}\texfamily
@@
 \\
-\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~\char95 )~=~{\itshape Nothing}
+\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~~~\char95 )~~~=~{\itshape Nothing}
 \\
-\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape TreeRight}~l~cxt)
+\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Right}~l~cxt)
 \end{tabbing}
-  \text{\ttfamily TreeContext} stores everything we need to reconstruct the hole, and \text{\ttfamily tzUp} does
+  \text{\texfamily {\itshape Context}} stores everything we need to reconstruct the hole, and \text{\texfamily up} does
   exactly that:
@@
 \\
-\texfamily ~~up~({\itshape Zipper}~r~({\itshape TreeRight}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
+\texfamily ~~up~({\itshape Zipper}~r~({\itshape Right}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
 \end{tabbing}
-  Implementation of \text{\ttfamily tzLeft}, \text{\ttfamily tzRight}, and \text{\ttfamily tzEnter} is very similar to the
-  List zipper case and is left as an exercise for the reader. \text{\ttfamily tzLeave} differs
+  Implementations of \text{\texfamily left}, \text{\texfamily right}, and \text{\texfamily enter} are very similar to the
+  list zipper case and are left as an exercise for the reader. \text{\texfamily leave} differs
   slightly in that we now move all the way up rather than left:
@@
 
-  The list and binary tree zipper we have considered here are homogeneous
-  zippers: the type of focus does not change upon zipper movement. Such zippers
-  can be a very useful abstraction. For example, a well-known window manager
-  XMonad\cite{xmonad} uses a rose tree zipper to track the window under focus.
-  For other tasks, however, one might need to traverse heterogeneous structures.
-  A zipper that can accomodate such needs is usually called a generic zipper as
-  it relies only on the generic structure of Algebraic Data Types (ADTs). One
-  can view an ADT as an Abstract Syntax Tree (AST) where each node is a Haskell
-  constructor rather than a syntax construct.
+  \paragraph{Generic Zipper} The list and binary tree zippers we have considered
+  so far are homogeneous zippers: the type of the focus does not change upon zipper
+  movement. Such a zipper can be a very useful abstraction. For example, a
+  well-known window manager XMonad\cite{xmonad} uses a rose tree zipper to track
+  the window under focus. For other tasks, however, one might need to traverse
+  heterogeneous structures. A zipper that can accomodate such needs is usually
+  called a \textit{generic zipper} as it relies only on the generic structure of
+  Algebraic Data Types (ADTs). One can view an ADT as an Abstract Syntax Tree
+  (AST) where each node is a Haskell constructor rather than a syntax construct.
 
@@
   \textit{dissectible}, i.e. we would like to be able to dissect the value into
-  the constructor and its arguments. Even though \text{\ttfamily Data\char46{}Data\char46{}gfold} allows us to achieve
-  this, we define out own typeclass which additionally allows us to propagate
-  down arbitrary constraints:
+  the constructor and its arguments. Even though \text{\texfamily {\itshape Data}.{\itshape Data}.gfoldl} allows us to
+  achieve this, we define out own typeclass which additionally allows us to
+  propagate down arbitrary constraints:
 \begin{tabbing}\texfamily
@@
 \\
-\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
+\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)
+\\
+\texfamily ~~~~~~~~~~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
 \end{tabbing}
-  For example, here is how we can make \text{\ttfamily Tree} an instance of \text{\ttfamily Disssectible}
+  For example, here is how we can make \text{\texfamily {\itshape Tree}} an instance of \text{\texfamily {\itshape Disssectible}}
 \begin{tabbing}\texfamily
@@
 \end{tabbing}
-  We can unpack a \text{\ttfamily Fork} and the zipper will thus be able to go down. \text{\ttfamily Leaf}s,
-  however, are left untouched and trying to go down from a \text{\ttfamily Leaf} will return
-  \text{\ttfamily Nothing}.
+  We can unpack a \text{\texfamily {\itshape Fork}} and the zipper will thus be able to go down. \text{\texfamily {\itshape Leaf}}s,
+  however, are left untouched and trying to go down from a \text{\texfamily {\itshape Leaf}} will return
+  \text{\texfamily {\itshape Nothing}}.
 
@@
 \\
-\texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~b~\WithMath{->}{\rightarrow}~{\itshape Right}~c~provides~r~\WithMath{->}{\rightarrow}~{\itshape Right}~c~(b~\WithMath{->}{\rightarrow}~provides)~r
+\texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)
+\\
+\texfamily ~~~~~~~~~~=>~b~\WithMath{->}{\rightarrow}~{\itshape Right}~c~provides~r~\WithMath{->}{\rightarrow}~{\itshape Right}~c~(b~\WithMath{->}{\rightarrow}~provides)~r
 \end{tabbing}
-  For example, for a tuple \text{\ttfamily \char40{}Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char41{}}, we can have
+  For example, for a tuple \text{\texfamily ({\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int})}, we can have
 \begin{tabbing}\texfamily
@@
 \\
-\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\rmfamily\enskip\{-\# UNPACK  \#-\}\enskip}~!({\itshape LocalContext}~c~hole~rights~parent)
+\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~!({\itshape LocalContext}~c~hole~rights~parent)
 \\
@@
 \end{tabbing}
-  And just like before the \text{\ttfamily Zipper} is a product of the hole and context:
+  And just like before the \text{\texfamily {\itshape Zipper}} is a product of the hole and context:
 \begin{tabbing}\texfamily
@@
   Implementation of movements is quite straightforward and is left out. Please,
-  refer to (\TODO{github repo}) for complete code.
-
-  We now consider a rather interesting application of generic zipper: embedding of
-  attribute grammars.
+  refer to (\TODO{github repo}) for complete code. We now consider a rather
+  interesting application of generic zipper: embedding of attribute grammars.
 
@@
 
-  Let us consider the repmin\footnote{%
-    The repmin problem:
+  Let us consider the repmin problem as an example of a problem that requires
+  multiple traversals:
   \begin{displayquote}
@@
   \end{displayquote}
-  } problem as an example of a problem that requires multiple traversals. The
-  classical solution is the following circular program:
+  The classical solution is the following circular program:
 \begin{tabbing}\texfamily

### 2018-11-20 doc/Paper.tex
Update PDF to d22bf7b3c88c727408f851664c85bea5ffb5f91a

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{\TODO{What's the title?}}
+\title{Modern type-safe embedding of attribute grammars}
+\subtitle{An exercise with functional zippers in Haskell}
 
@@
 \institute{%
-  LISP/Release -- Universidade da Beira Interior, Portugal \email{jpf@di.ubi.pt} \and
+  CISUC -- Universidade de Coimbra, Portugal \email{jpf@dei.uc.pt} \and
   University of California, Irvine, USA \email{pribeiro@uci.edu} \and
-  Universidad de la  Rep\'{u}blica, Uruguay \email{\{pardo,mviera\}@fing.edu.uy} \and
-  Universidade do Minho, Portugal \email{saraiva@di.uminho.pt} \and
-  Radboud University, The Netherlands \email{twesterhout@student.ru.nl}%
+  Universidad de la  Rep\'{u}blica, Uruguay, \email{\{pardo,mviera\}@fing.edu.uy} \and
+  Universidade do Minho, Portugal, \email{saraiva@di.uminho.pt} \and
+  Radboud University, The Netherlands, \email{twesterhout@student.ru.nl}%
 }
@@
 \begin{abstract}
-  \TODO{What's the abstract?}
+
+  % Introduction. In one sentence, what’s the topic?
+
+  Attribute grammars are a powerful, declarative formalism to implement and
+  reason about programs which, by design, are conveniently modular.
+
+  % State the problem you tackle
+
+  Although a full attribute grammar compiler can be tailored to specific needs,
+  its implementation is highly non trivial, and its long term maintenance is a
+  major endeavor.
+
+  % Summarize (in one sentence) why nobody else has adequately answered the
+  % research question yet.
+
+  In fact, maintaining a traditional attribute grammar system is such a hard
+  effort that most such system that were proposed in the past are no longer
+  active.
+
+  % Explain, in one sentence, how you tackled the research question.
+
+  Our approach to implement attribute grammars is to write them as first class
+  citizens of a modern functional programming language.
+
+  % In one sentence, how did you go about doing the research that follows from
+  % your big idea.
+
+  We improve a previous zipped-based attribute grammar embedding making it
+  non-intrusive (i.e. no changes need to be made to the user-defined data types)
+  and type-safe. On top of that, we achieve clearer syntax by using modern
+  Haskell extensions.
+
+  % As a single sentence, what’s the key impact of your research?
+
+  We believe our embedding can be employed in practice to implement elegant,
+  efficient and modular solutions to real life programming challenges.
 
@@
 
+  Attribute Grammars (\ags) are a declarative formalism that was proposed by
+  Knuth~\cite{Knuth68} in the late 60s and allows the implementation and
+  reasoning about programs in a modular and convenient way. A concrete \ag
+  relies on a context-free grammar to define the syntax of a language, and on
+  attributes associated to the productions of the grammar to define the
+  semantics of that language.
+  %while adding \emph{attributes} to it so that it is also possible to define
+  %its semantics.
+  \ags\ have been used in practice to specify real programming languages, like
+  for example \HASKELL~\cite{DijkstraFS09}, as well as powerful pretty printing
+  algorithms \cite{SPS99}, deforestation techniques \cite{joao07pepm} and
+  powerful type systems \cite{MiddelkoopDS10}.
+
+  When programming with \ags, modularity is achieved due to the possibility of
+  defining and using different aspects of computations as separate attributes.
+  Attributes are distinct computation units, typically quite simple and modular,
+  that can be combined into elaborated solutions to complex programming
+  problems. They can also be analyzed, debugged and maintained independently
+  which eases program development and evolution.
+
+  \ags\ have proven to be particularly useful to specify computations over
+  trees: given one tree, several \ag\ systems such
+  as~\cite{syngen,uuag,lrc,silver} take specifications of which values, or
+  attributes, need to be computed on the tree and perform these computations.
+  The design and coding efforts put into the creation, improvement and
+  maintenance of these \ag\ systems, however, is tremendous, which often is an
+  obstacle to achieving the success they deserve.
+
+  An increasingly popular alternative approach to the use of \ags\ relies on
+  embedding them as first class citizens of general purpose programming
+  languages~\cite{Oege00,DBLP:conf/sblp/MartinsFS13,erlangAGs,kiama,doaitse09icfp,balestrieri}.
+  This avoids the burden of implementing a totally new language and associated
+  system by hosting it in state-of-the-art programming languages. Following this
+  approach one then exploits the modern constructions and infrastructure that
+  are already provided by those languages and focus on the particularities of
+  the domain specific language being developed.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{tabbing}\texfamily
-~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]
-\\
-\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0
-\\
-\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j
-\\
-\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after
-\\
-\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)
-\\
+~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]\\
+\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0\\
+\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j\\
+\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after\\
+\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)\\
 \texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
@@
 \begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
-\\
+~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 \\
 \texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
@@
 \begin{tabbing}\texfamily
-~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}
-\\
-\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$
-\\
-\texfamily ~~~~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
-\\
-\texfamily 
-\\
-\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}
-\\
-\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$
-\\
+~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}\\
+\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$\\
+\texfamily ~~~~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)\\
+\texfamily \\
+\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}\\
+\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$\\
 \texfamily ~~~~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
@@
 \begin{tabbing}\texfamily
-~~enter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~enter~[]~~~~~~~=~{\itshape Nothing}
-\\
-\texfamily ~~enter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
-\\
-\texfamily 
-\\
-\texfamily ~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
-\\
+~~enter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~enter~[]~~~~~~~=~{\itshape Nothing}\\
+\texfamily ~~enter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)\\
+\texfamily \\
+\texfamily ~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]\\
 \texfamily ~~leave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
@@
 \begin{tabbing}\texfamily
-~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~modify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
-\\
-\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))
-\\
-\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}
-\\
-\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')
-\\
-\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}
-\\
-\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)
-\\
+~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~modify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}\\
+\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))\\
+\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}\\
+\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')\\
+\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}\\
+\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)\\
 \texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
@@
 \begin{tabbing}\texfamily
-~~modifyExample~::~{\itshape IO}~()
-\\
-\texfamily ~~modifyExample~=~print~\$
-\\
-\texfamily ~~~~enter~@{\itshape Int}~>=>~right
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~right
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(const~[])
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~(+1))
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~left
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~negate)
-\\
+~~modifyExample~::~{\itshape IO}~()\\
+\texfamily ~~modifyExample~=~print~\$\\
+\texfamily ~~~~enter~@{\itshape Int}~>=>~right\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~right\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(const~[])\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~(+1))\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~left\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~negate)\\
 \texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~leave~\$~[1,~2,~3,~4,~5]
@@
 \begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~{\itshape Tree}~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
-\\
-\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Top}
-\\
-\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
-\\
+~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~{\itshape Tree}~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 \\
+\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Top}\\
+\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)\\
 \texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Right}~({\itshape Tree}~a)~!({\itshape Context}~a)
@@
 \begin{tabbing}\texfamily
-~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~~~\char95 )~~~=~{\itshape Nothing}
-\\
+~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~~~\char95 )~~~=~{\itshape Nothing}\\
 \texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Right}~l~cxt)
@@
 \begin{tabbing}\texfamily
-~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}
-\\
-\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
-\\
+~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}\\
+\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt\\
 \texfamily ~~up~({\itshape Zipper}~r~({\itshape Right}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
@@
 \begin{tabbing}\texfamily
-~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a
-\\
-\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}
-\\
-\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'
-\\
+~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a\\
+\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}\\
+\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'\\
 \texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
@@
 \begin{tabbing}\texfamily
-~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}
-\\
-\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a
-\\
-\texfamily 
-\\
-\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}
-\\
-\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b
-\\
-\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)
-\\
+~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}\\
+\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a\\
+\texfamily \\
+\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}\\
+\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b\\
+\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)\\
 \texfamily ~~~~~~~~~~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
@@
 \begin{tabbing}\texfamily
-~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}
-\\
-\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r
-\\
+~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}\\
+\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r\\
 \texfamily ~~~~dissect~x~=~{\itshape LOne}~x
@@
 \begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Right}~c~provides~r~{\bfseries where}
-\\
-\texfamily ~~~~{\itshape RNil}~~::~{\itshape Right}~c~r~r
-\\
-\texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)
-\\
+~~{\bfseries data}~{\itshape Right}~c~provides~r~{\bfseries where}\\
+\texfamily ~~~~{\itshape RNil}~~::~{\itshape Right}~c~r~r\\
+\texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)\\
 \texfamily ~~~~~~~~~~=>~b~\WithMath{->}{\rightarrow}~{\itshape Right}~c~provides~r~\WithMath{->}{\rightarrow}~{\itshape Right}~c~(b~\WithMath{->}{\rightarrow}~provides)~r
@@
 \begin{tabbing}\texfamily
-~~lefts~=~{\itshape LOne}~(,,,,,)~`{\itshape LCons}`~1~`{\itshape LCons}`~2~`{\itshape LCons}`~3
-\\
-\texfamily ~~hole~=~4
-\\
+~~lefts~=~{\itshape LOne}~(,,,,,)~`{\itshape LCons}`~1~`{\itshape LCons}`~2~`{\itshape LCons}`~3\\
+\texfamily ~~hole~=~4\\
 \texfamily ~~rights~=~5~`{\itshape RCons}`~6~`{\itshape RCons}`~{\itshape RNil}
@@
 \begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape LocalContext}~c~hole~rights~parent~=
-\\
-\texfamily ~~~~{\itshape LocalContext}~!({\itshape Left}~c~(hole~\WithMath{->}{\rightarrow}~rights))~!({\itshape Right}~c~rights~parent)
-\\
-\texfamily 
-\\
-\texfamily ~~{\bfseries data}~{\itshape Context}~::~({\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}
-\\
-\texfamily ~~~~{\itshape RootContext}~::~\char'24~c~root.~{\itshape Context}~c~root~root
-\\
-\texfamily ~~~~(:>)~::~\char'24~c~parent~root~hole~rights.~(c~parent,~{\itshape Dissectible}~c~parent)
-\\
-\texfamily ~~~~~~~~~=>~!({\itshape Context}~c~parent~root)
-\\
-\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~!({\itshape LocalContext}~c~hole~rights~parent)
-\\
+~~{\bfseries data}~{\itshape LocalContext}~c~hole~rights~parent~=\\
+\texfamily ~~~~{\itshape LocalContext}~!({\itshape Left}~c~(hole~\WithMath{->}{\rightarrow}~rights))~!({\itshape Right}~c~rights~parent)\\
+\texfamily \\
+\texfamily ~~{\bfseries data}~{\itshape Context}~::~({\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}\\
+\texfamily ~~~~{\itshape RootContext}~::~\char'24~c~root.~{\itshape Context}~c~root~root\\
+\texfamily ~~~~(:>)~::~\char'24~c~parent~root~hole~rights.~(c~parent,~{\itshape Dissectible}~c~parent)\\
+\texfamily ~~~~~~~~~=>~!({\itshape Context}~c~parent~root)\\
+\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~!({\itshape LocalContext}~c~hole~rights~parent)\\
 \texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Context}~c~hole~root
@@
 \begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Zipper}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(root~::~{\itshape Type})~=
-\\
-\texfamily ~~~~\char'24~hole.~(c~hole,~{\itshape Dissectible}~c~hole)~=>
-\\
-\texfamily ~~~~~~{\itshape Zipper}~\char123 ~\char95 zHole~::~!hole
-\\
-\texfamily ~~~~~~~~~~~~~,~\char95 zCxt~~::~!({\itshape Context}~c~hole~root)
-\\
+~~{\bfseries data}~{\itshape Zipper}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(root~::~{\itshape Type})~=\\
+\texfamily ~~~~\char'24~hole.~(c~hole,~{\itshape Dissectible}~c~hole)~=>\\
+\texfamily ~~~~~~{\itshape Zipper}~\char123 ~\char95 zHole~::~!hole\\
+\texfamily ~~~~~~~~~~~~~,~\char95 zCxt~~::~!({\itshape Context}~c~hole~root)\\
 \texfamily ~~~~~~~~~~~~~\char125 
@@
 \begin{tabbing}\texfamily
-~~repmin~::~{\itshape Tree}~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Tree}~{\itshape Int}
-\\
-\texfamily ~~repmin~t~=~t'
-\\
-\texfamily ~~~~{\bfseries where}~(t',~m')~=~go~t~m'
-\\
-\texfamily ~~~~~~~~~~go~({\itshape Leaf}~x~~~~)~m~=~({\itshape Leaf}~m,~x)
-\\
-\texfamily ~~~~~~~~~~go~({\itshape Fork}~xs~ys)~m~=~({\itshape Fork}~xs'~ys',~min~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$}~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})
-\\
-\texfamily ~~~~~~~~~~~~{\bfseries where}~(xs',~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$})~=~go~xs~m
-\\
+~~repmin~::~{\itshape Tree}~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Tree}~{\itshape Int}\\
+\texfamily ~~repmin~t~=~t'\\
+\texfamily ~~~~{\bfseries where}~(t',~m')~=~go~t~m'\\
+\texfamily ~~~~~~~~~~go~({\itshape Leaf}~x~~~~)~m~=~({\itshape Leaf}~m,~x)\\
+\texfamily ~~~~~~~~~~go~({\itshape Fork}~xs~ys)~m~=~({\itshape Fork}~xs'~ys',~min~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$}~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})\\
+\texfamily ~~~~~~~~~~~~{\bfseries where}~(xs',~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$})~=~go~xs~m\\
 \texfamily ~~~~~~~~~~~~~~~~~~(ys',~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})~=~go~ys~m
@@
 \begin{tabbing}\texfamily
-~~SYN~{\itshape Tree}~{\itshape Int}~[localMin~:~{\itshape Int}]
-\\
-\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.localMin~=~@value
-\\
-\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~min~@left.localMin
-\\
-\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.localMin
-\\
-\texfamily 
-\\
-\texfamily ~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]
-\\
-\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin
-\\
-\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated
-\\
-\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.updated
-\\
-\texfamily 
-\\
-\texfamily 
-\\
-\texfamily ~~INH~{\itshape Tree}~{\itshape Int}~[~globalMin~:~{\itshape Int}~]
-\\
-\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Fork}~left.globalMin~~=~@lhs.globalMin
-\\
-\texfamily ~~~~~~~~~~~~~~~~~~~~~~right.globalMin~=~@lhs.globalMin
-\\
-\texfamily 
-\\
-\texfamily ~~DATA~{\itshape Root}~|~{\itshape Root}~tree~:~{\itshape Tree}~{\itshape Int}
-\\
+~~SYN~{\itshape Tree}~{\itshape Int}~[localMin~:~{\itshape Int}]\\
+\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.localMin~=~@value\\
+\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~min~@left.localMin\\
+\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.localMin\\
+\texfamily \\
+\texfamily ~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]\\
+\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin\\
+\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated\\
+\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.updated\\
+\texfamily \\
+\texfamily \\
+\texfamily ~~INH~{\itshape Tree}~{\itshape Int}~[~globalMin~:~{\itshape Int}~]\\
+\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Fork}~left.globalMin~~=~@lhs.globalMin\\
+\texfamily ~~~~~~~~~~~~~~~~~~~~~~right.globalMin~=~@lhs.globalMin\\
+\texfamily \\
+\texfamily ~~DATA~{\itshape Root}~|~{\itshape Root}~tree~:~{\itshape Tree}~{\itshape Int}\\
 \texfamily ~~SEM~~{\itshape Root}~|~{\itshape Root}~tree.globalMin~=~@tree.localMin
@@
 \begin{tabbing}\texfamily
-~~localMin~::~{\itshape Zipper}~({\itshape WhereAmI}~{\itshape Position})~({\itshape Tree}~{\itshape Int})~\WithMath{->}{\rightarrow}~{\itshape Int}
-\\
-\texfamily ~~localMin~z@({\itshape Zipper}~hole~\char95 )~=~{\bfseries case}~whereami~hole~{\bfseries of}
-\\
-\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Leaf}~x~=~hole~{\bfseries in}~x
-\\
-\texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Just}~l~=~child~0~z;~{\itshape Just}~r~=~child~1~z
-\\
+~~localMin~::~{\itshape Zipper}~({\itshape WhereAmI}~{\itshape Position})~({\itshape Tree}~{\itshape Int})~\WithMath{->}{\rightarrow}~{\itshape Int}\\
+\texfamily ~~localMin~z@({\itshape Zipper}~hole~\char95 )~=~{\bfseries case}~whereami~hole~{\bfseries of}\\
+\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Leaf}~x~=~hole~{\bfseries in}~x\\
+\texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Just}~l~=~child~0~z;~{\itshape Just}~r~=~child~1~z\\
 \texfamily ~~~~~~~~~~~~~~{\bfseries in}~min~(localMin~l)~(localMin~r)
@@
 \begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Position}~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}
-\\
-\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})
-\\
+~~{\bfseries data}~{\itshape Position}~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}\\
+\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})\\
 \texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})

### 2018-11-20 doc/Paper.tex
Update PDF to 2b097bdc74302c95fa6ee9c4ed4d6e8f80c7caf2

@@
 
-  Attribute Grammars (\ags) are a declarative formalism that was proposed by
+  Attribute Grammars (AGs) are a declarative formalism that was proposed by
   Knuth~\cite{Knuth68} in the late 60s and allows the implementation and
-  reasoning about programs in a modular and convenient way. A concrete \ag
+  reasoning about programs in a modular and convenient way. A concrete AG
   relies on a context-free grammar to define the syntax of a language, and on
@@
   %its semantics.
-  \ags\ have been used in practice to specify real programming languages, like
-  for example \HASKELL~\cite{DijkstraFS09}, as well as powerful pretty printing
+  AGs have been used in practice to specify real programming languages, like
+  for example Haskell \cite{DijkstraFS09}, as well as powerful pretty printing
   algorithms \cite{SPS99}, deforestation techniques \cite{joao07pepm} and
@@
 
-  When programming with \ags, modularity is achieved due to the possibility of
+  When programming with AGs, modularity is achieved due to the possibility of
   defining and using different aspects of computations as separate attributes.
@@
 
-  \ags\ have proven to be particularly useful to specify computations over
-  trees: given one tree, several \ag\ systems such
+  AGs have proven to be particularly useful to specify computations over
+  trees: given one tree, several AG systems such
   as~\cite{syngen,uuag,lrc,silver} take specifications of which values, or
@@
   The design and coding efforts put into the creation, improvement and
-  maintenance of these \ag\ systems, however, is tremendous, which often is an
+  maintenance of these AG systems, however, is tremendous, which often is an
   obstacle to achieving the success they deserve.
 
-  An increasingly popular alternative approach to the use of \ags\ relies on
+  An increasingly popular alternative approach to the use of AGs relies on
   embedding them as first class citizens of general purpose programming
@@
   the domain specific language being developed.
+
+  Functional zipper\cite{huet1997zipper} is a powerful abstraction which greatly
+  simplifies the implementation of traversal algorithms performing a lot of
+  local updates. Functional zippers have successfully been applied to constuct
+  an attribute grammar embeddings in Haskell\cite{...}. Despite its elegance,
+  this solution had a major drawback which prevented its use in real-world
+  applications: attributes were not cached, but rather repeatedly recomputed
+  which severely hurt performance. Recently, this flaw has been
+  eliminated\cite{...} and replaced with a different one: the approach became
+  intrusive, i.e. to benefit from the embedding user-defined data structures
+  have to be adjusted.
+
+  In this paper we present an alternative mechanism to cache attributes based on
+  a self-organising infinite grid. This graph is laid on top of the user-defined
+  algebraic data type and mirrors its structure. The used-defined data type
+  remains untouched. The embedding is then based on two coherent zippers (rather
+  than one) traversing the data structures in parallel. On top of being
+  non-intrusive our solution is completely type-safe. Modern Haskell extensions
+  such as \texttt{ConstraintKinds} allow us to propagate constraints down in the
+  ADT completely eliminating run-time type casts present in the previous
+  versions. Another side benefit of using modern Haskell is a cleaner syntax
+  with less code being generated with Template Haskell.
 

### 2018-11-21 doc/Paper.tex
Update PDF to 8e3fa25d51471df5f8313ae186ee450073dac8cd

--- 2018-11-20 doc/Paper.tex
@@
 \begin{abstract}
-  \TODO{What's the abstract?}
+  % Introduction. In one sentence, what’s the topic?
+
+  Attribute grammars are a powerful, declarative formalism to implement and
+  reason about programs which, by design, are conveniently modular.
+
+  % State the problem you tackle
+
+  Although a full attribute grammar compiler can be tailored to specific needs,
+  its implementation is highly non trivial, and its long term maintenance is a
+  major endeavor.
+
+  % Summarize (in one sentence) why nobody else has adequately answered the
+  % research question yet.
+
+  In fact, maintaining a traditional attribute grammar system is such a hard
+  effort that most such systems that were proposed in the past are no longer
+  active.
+
+  % Explain, in one sentence, how you tackled the research question.
+
+  Our approach to implement attribute grammars is to write them as first class
+  citizens of a modern functional programming language.
+
+  % In one sentence, how did you go about doing the research that follows from
+  % your big idea.
+
+  We improve a previous zipped-based attribute grammar embedding making it
+  non-intrusive (i.e. no changes need to be made to the user-defined data types)
+  and type-safe. On top of that, we achieve clearer syntax by using modern
+  Haskell extensions.
+
+  % As a single sentence, what’s the key impact of your research?
+
+  We believe our embedding can be employed in practice to implement elegant,
+  efficient and modular solutions to real life programming challenges.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Functional Zippers}
-  Zipper is a data structure commonly used in functional programming for
-  traversal with fast local updates. The zipper data structure was originally
-  conceived by Huet\cite{huet1997zipper} in the context of trees. We will,
-  however, first consider a simpler problem: a bidirectional list traversal.
-
-  \paragraph{Lists} Suppose that we would like to update a list at a specific
-  position:
-\begin{tabbing}\texfamily
-~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]
-\\
-\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0
-\\
-\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j
-\\
-\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after
-\\
-\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)
-\\
-\texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
-\end{tabbing}
-  Here \text{\texfamily modify} takes an update action \text{\texfamily f}\footnote{\text{\texfamily f} returns a list rather
-  than a single element to prevent curious readers from suggesting to use a
-  boxed array instead of a list.}, an index \text{\texfamily i}, and a list \text{\texfamily xs} and returns a
-  new list with the \text{\texfamily i}'th element replaced with the result of \text{\texfamily f}.%
-  First, we ``unpack'' the list into $\text{\texfamily before\Sp =\Sp [}\ttsub{\text{\texfamily xs}}{\text{\texfamily 0}},\dots
-  \ttsub{\text{\texfamily xs}}{\text{\texfamily i-1}}\text{\texfamily ]}$, $\text{\texfamily x\Sp =\Sp }\ttsub{\text{\texfamily xs}}{\text{\texfamily i}}$, and
-  $\text{\texfamily after\Sp =\Sp [}\ttsub{\text{\texfamily xs}}{\text{\texfamily i+1}},\dots\text{\texfamily ]}$. Then replace \text{\texfamily x} by \text{\texfamily f(x)} and
-  finally ``pack'' the result back into a single list. If we do a lot of
-  updates, we end up unpacking and packing the list over and over again -- very
-  time-consuming for long lists. We would thus like to benefit from fusion
-  without explicitly working with the unpacked representation as it is
-  bug-prone. A list zipper provides a way to achieve this.
-
-  A zipper consists of a focus (alternatively called a hole) and surrounding
-  context\footnote{`\text{\texfamily !}' in front of the type of \text{\texfamily \char95 cxt} is a
-  strictness annotation enabled by the \text{\ttfamily BangPatterns} extension. Context can be
-  seen as a path from the to the current focus and is always finite, i.e. there
-  is no reason for it to be lazy.}:
-\begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
-\\
-\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
-\end{tabbing}
-  where the \text{\texfamily {\itshape Context}} keeps track of elements to the left (\text{\texfamily before} in
-  \text{\texfamily modify}) and to the right (\text{\texfamily after} in \text{\texfamily modify}) of the focus. We can now
-  define movements:
-\begin{tabbing}\texfamily
-~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}
-\\
-\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$
-\\
-\texfamily ~~~~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
-\\
-\texfamily 
-\\
-\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}
-\\
-\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$
-\\
-\texfamily ~~~~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
-\end{tabbing}
-  and functions for entering and leaving the zipper:
-\begin{tabbing}\texfamily
-~~enter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~enter~[]~~~~~~~=~{\itshape Nothing}
-\\
-\texfamily ~~enter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
-\\
-\texfamily 
-\\
-\texfamily ~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
-\\
-\texfamily ~~leave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
-\end{tabbing}
-
-  Finally, we define a local version of our \text{\texfamily modify} function (\TODO{Boy, is
-  this function ugly...})
-\begin{tabbing}\texfamily
-~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~modify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
-\\
-\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))
-\\
-\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}
-\\
-\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')
-\\
-\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}
-\\
-\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)
-\\
-\texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
-\end{tabbing}
-  using which we can perform multiple updates efficiently and with minimal code
-  bloat\footnote{%
-  Operator \text{\ttfamily \char62{}\char61{}\char62{}} comes from \text{\texfamily {\itshape {\itshape Control}.Monad}} module in \text{\ttfamily base}
-  and has the following signature:%
-\begin{tabbing}\texfamily
-~~(>=>)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
-\end{tabbing}
-%
-  }:
-\begin{tabbing}\texfamily
-~~modifyExample~::~{\itshape IO}~()
-\\
-\texfamily ~~modifyExample~=~print~\$
-\\
-\texfamily ~~~~enter~@{\itshape Int}~>=>~right
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~right
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(const~[])
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~(+1))
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~left
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~negate)
-\\
-\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~leave~\$~[1,~2,~3,~4,~5]
-\end{tabbing}
-
-  \paragraph{Trees} Consider now a binary tree data structure:
-\begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Tree}~a~=~{\itshape Leaf}~!a~|~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
-\end{tabbing}
-  A binary tree zipper is slightly more insteresting than the list zipper,
-  because we can move up and down the tree as well as left and right. The zipper
-  again consists of a hole (a subtree we are focused on) and its surrounding
-  context (a path from the hole to the root of the tree):
-\begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~{\itshape Tree}~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
-\\
-\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Top}
-\\
-\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
-\\
-\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Right}~({\itshape Tree}~a)~!({\itshape Context}~a)
-\end{tabbing}
-  To move the zipper down, we ``unpack'' the current hole:
-\begin{tabbing}\texfamily
-~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~~~\char95 )~~~=~{\itshape Nothing}
-\\
-\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Right}~l~cxt)
-\end{tabbing}
-  \text{\texfamily {\itshape Context}} stores everything we need to reconstruct the hole, and \text{\texfamily up} does
-  exactly that:
-\begin{tabbing}\texfamily
-~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
-\\
-\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}
-\\
-\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
-\\
-\texfamily ~~up~({\itshape Zipper}~r~({\itshape Right}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
-\end{tabbing}
-  Implementations of \text{\texfamily left}, \text{\texfamily right}, and \text{\texfamily enter} are very similar to the
-  list zipper case and are left as an exercise for the reader. \text{\texfamily leave} differs
-  slightly in that we now move all the way up rather than left:
-\begin{tabbing}\texfamily
-~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a
-\\
-\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}
-\\
-\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'
-\\
-\texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
-\end{tabbing}
-
-  \paragraph{Generic Zipper} The list and binary tree zippers we have considered
-  so far are homogeneous zippers: the type of the focus does not change upon zipper
-  movement. Such a zipper can be a very useful abstraction. For example, a
-  well-known window manager XMonad\cite{xmonad} uses a rose tree zipper to track
-  the window under focus. For other tasks, however, one might need to traverse
-  heterogeneous structures. A zipper that can accomodate such needs is usually
-  called a \textit{generic zipper} as it relies only on the generic structure of
-  Algebraic Data Types (ADTs). One can view an ADT as an Abstract Syntax Tree
-  (AST) where each node is a Haskell constructor rather than a syntax construct.
-
-  The generic zipper we will use is very similar to the one presented
-  in\cite{adams2010syz}. The most common technique in Haskell for supporting
-  heterogeneous types is Existential Quantification. However, not every type can
-  act as a hole. To support moving down the tree, we need the hole to be
-  \textit{dissectible}, i.e. we would like to be able to dissect the value into
-  the constructor and its arguments. Even though \text{\texfamily {\itshape Data}.{\itshape Data}.gfoldl} allows us to
-  achieve this, we define out own typeclass which additionally allows us to
-  propagate down arbitrary constraints:
-\begin{tabbing}\texfamily
-~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}
-\\
-\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a
-\\
-\texfamily 
-\\
-\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}
-\\
-\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b
-\\
-\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)
-\\
-\texfamily ~~~~~~~~~~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
-\end{tabbing}
-  For example, here is how we can make \text{\texfamily {\itshape Tree}} an instance of \text{\texfamily {\itshape Disssectible}}
-\begin{tabbing}\texfamily
-~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}
-\\
-\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r
-\\
-\texfamily ~~~~dissect~x~=~{\itshape LOne}~x
-\end{tabbing}
-  We can unpack a \text{\texfamily {\itshape Fork}} and the zipper will thus be able to go down. \text{\texfamily {\itshape Leaf}}s,
-  however, are left untouched and trying to go down from a \text{\texfamily {\itshape Leaf}} will return
-  \text{\texfamily {\itshape Nothing}}.
-
-  To allow the zipper to move left and right, we need a means to encode
-  arguments to the right of the hole. Following Adams et al, we define a GADT
-  representing constructor arguments to the right of the hole:
-\begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Right}~c~provides~r~{\bfseries where}
-\\
-\texfamily ~~~~{\itshape RNil}~~::~{\itshape Right}~c~r~r
-\\
-\texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)
-\\
-\texfamily ~~~~~~~~~~=>~b~\WithMath{->}{\rightarrow}~{\itshape Right}~c~provides~r~\WithMath{->}{\rightarrow}~{\itshape Right}~c~(b~\WithMath{->}{\rightarrow}~provides)~r
-\end{tabbing}
-  For example, for a tuple \text{\texfamily ({\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int})}, we can have
-\begin{tabbing}\texfamily
-~~lefts~=~{\itshape LOne}~(,,,,,)~`{\itshape LCons}`~1~`{\itshape LCons}`~2~`{\itshape LCons}`~3
-\\
-\texfamily ~~hole~=~4
-\\
-\texfamily ~~rights~=~5~`{\itshape RCons}`~6~`{\itshape RCons}`~{\itshape RNil}
-\end{tabbing}
-  generalising this a little, we arrive at:
-\begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape LocalContext}~c~hole~rights~parent~=
-\\
-\texfamily ~~~~{\itshape LocalContext}~!({\itshape Left}~c~(hole~\WithMath{->}{\rightarrow}~rights))~!({\itshape Right}~c~rights~parent)
-\\
-\texfamily 
-\\
-\texfamily ~~{\bfseries data}~{\itshape Context}~::~({\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}
-\\
-\texfamily ~~~~{\itshape RootContext}~::~\char'24~c~root.~{\itshape Context}~c~root~root
-\\
-\texfamily ~~~~(:>)~::~\char'24~c~parent~root~hole~rights.~(c~parent,~{\itshape Dissectible}~c~parent)
-\\
-\texfamily ~~~~~~~~~=>~!({\itshape Context}~c~parent~root)
-\\
-\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~!({\itshape LocalContext}~c~hole~rights~parent)
-\\
-\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Context}~c~hole~root
-\end{tabbing}
-  And just like before the \text{\texfamily {\itshape Zipper}} is a product of the hole and context:
-\begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Zipper}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(root~::~{\itshape Type})~=
-\\
-\texfamily ~~~~\char'24~hole.~(c~hole,~{\itshape Dissectible}~c~hole)~=>
-\\
-\texfamily ~~~~~~{\itshape Zipper}~\char123 ~\char95 zHole~::~!hole
-\\
-\texfamily ~~~~~~~~~~~~~,~\char95 zCxt~~::~!({\itshape Context}~c~hole~root)
-\\
-\texfamily ~~~~~~~~~~~~~\char125 
-\end{tabbing}
-
-  Implementation of movements is quite straightforward and is left out. Please,
-  refer to (\TODO{github repo}) for complete code. We now consider a rather
-  interesting application of generic zipper: embedding of attribute grammars.
-
-\section{Attribute Grammars}
-  Attribute grammars (AGs) are an extension of context-free grammars that allow
-  to specify context-sensitive syntax as well as the semantics. AGs achieve it
-  by associating a set of attributes with each grammar symbol. These attributes
-  are defined using evaluation rules assiciated with production rules of the
-  context-free grammar.
-
-  Attributes are then usually divided into two disjoint sets: synthesized
-  attributes and the inherited attributes. Such distinction is required for the
-  construction of a dependency graph. It is then used for specification of the
-  evaluation order and detection of circularity. In the zipper-based embedding of
-  attribute grammars we make no use of a dependency graph and thus do not divide
-  attributes into classes.
-
-  Let us consider the repmin problem as an example of a problem that requires
-  multiple traversals:
-  \begin{displayquote}
-    Given a tree of integers, replace every integer with the
-    minimum integer in the tree, in one pass.
-  \end{displayquote}
-  The classical solution is the following circular program:
-\begin{tabbing}\texfamily
-~~repmin~::~{\itshape Tree}~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Tree}~{\itshape Int}
-\\
-\texfamily ~~repmin~t~=~t'
-\\
-\texfamily ~~~~{\bfseries where}~(t',~m')~=~go~t~m'
-\\
-\texfamily ~~~~~~~~~~go~({\itshape Leaf}~x~~~~)~m~=~({\itshape Leaf}~m,~x)
-\\
-\texfamily ~~~~~~~~~~go~({\itshape Fork}~xs~ys)~m~=~({\itshape Fork}~xs'~ys',~min~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$}~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})
-\\
-\texfamily ~~~~~~~~~~~~{\bfseries where}~(xs',~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$})~=~go~xs~m
-\\
-\texfamily ~~~~~~~~~~~~~~~~~~(ys',~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})~=~go~ys~m
-\end{tabbing}
-
-  Although quite elegant, the code lacks modularity and is very difficult to
-  reason about. Attribute Grammars provide a more modular approach. Viera et
-  al\cite{viera2009agsfly} identified three steps for solving repmin: computing
-  the minimal value, passing it down from the root to the leaves, and
-  constructing the resulting tree. We can associate each step with an
-  attribute\cite{swierstra1998designing}:
-  \begin{itemize}
-  \item A synthesized attribute \text{\texfamily localMin\Sp ::\Sp {\itshape Int}} represents the minimum value
-        of a subtree. Computing the minimal value thus corresponds to evaluation
-        of the \text{\texfamily localMin} attribute for the root tree.
-  \item An inherited attribute \text{\texfamily globalMin\Sp ::\Sp {\itshape Int}} is used to pass down the
-        minimal value.
-  \item Finally, a synthesized attribute \text{\texfamily updated\Sp ::\Sp {\itshape Tree}\Sp {\itshape Int}} is the subtree
-        with leaf values replaced by values of their \text{\texfamily globalMin} attributes. The
-        solution is thus the value of \text{\texfamily updated} attribute for the root tree.
-  \end{itemize}
-  The obtained AG is presented in figure~\ref{fig:repmin-AG}. \TODO{Do we
-  actually need to explain the algorithms here? It seems a little bit childish
-  to explain how to compute the minimum of a binary tree... If we absolutely
-  have to explain stuff, maybe just put it in the caption.}
-
-\begin{figure}
-\centering
-\fbox{\parbox{\linewidth}{%
-\begin{tabbing}\texfamily
-~~SYN~{\itshape Tree}~{\itshape Int}~[localMin~:~{\itshape Int}]
-\\
-\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.localMin~=~@value
-\\
-\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~min~@left.localMin
-\\
-\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.localMin
-\\
-\texfamily 
-\\
-\texfamily ~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]
-\\
-\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin
-\\
-\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated
-\\
-\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.updated
-\\
-\texfamily 
-\\
-\texfamily 
-\\
-\texfamily ~~INH~{\itshape Tree}~{\itshape Int}~[~globalMin~:~{\itshape Int}~]
-\\
-\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Fork}~left.globalMin~~=~@lhs.globalMin
-\\
-\texfamily ~~~~~~~~~~~~~~~~~~~~~~right.globalMin~=~@lhs.globalMin
-\\
-\texfamily 
-\\
-\texfamily ~~DATA~{\itshape Root}~|~{\itshape Root}~tree~:~{\itshape Tree}~{\itshape Int}
-\\
-\texfamily ~~SEM~~{\itshape Root}~|~{\itshape Root}~tree.globalMin~=~@tree.localMin
-\end{tabbing}
-%
-}}
-\caption{%
-  Attribute grammar for repmin. The syntax is closely mirrors the one used
-  in\cite{swierstra1998designing}. \text{\texfamily SYN} and \text{\texfamily INH} introduce synthesized
-  and inherited attributes respectively. \text{\texfamily SEM} is used for defining semantic
-  rules. A new data type \text{\texfamily {\itshape Root}} is introduced as it is common in the AG setting
-  to ``connect'' \text{\texfamily localMin} with \text{\texfamily globalMin}.
-}
-\label{fig:repmin-AG}
-\end{figure}
-
-  We now move on to embed this attribute grammar into Haskell. Semantic rules
-  simply become functions, which, given a zipper, return values of the
-  attributes. For example,
-\begin{tabbing}\texfamily
-~~localMin~::~{\itshape Zipper}~({\itshape WhereAmI}~{\itshape Position})~({\itshape Tree}~{\itshape Int})~\WithMath{->}{\rightarrow}~{\itshape Int}
-\\
-\texfamily ~~localMin~z@({\itshape Zipper}~hole~\char95 )~=~{\bfseries case}~whereami~hole~{\bfseries of}
-\\
-\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Leaf}~x~=~hole~{\bfseries in}~x
-\\
-\texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Just}~l~=~child~0~z;~{\itshape Just}~r~=~child~1~z
-\\
-\texfamily ~~~~~~~~~~~~~~{\bfseries in}~min~(localMin~l)~(localMin~r)
-\end{tabbing}
-
-  Apart from the type signature, the code is pretty straightforward and closely
-  mirrors the AG we defined earlier. \text{\texfamily whereami} function allows us to ``look
-  around'' and returns the position of the zipper. Thus the \text{\texfamily {\bfseries case}} corresponds
-  to the pattern matches on the left of the vertical bars on
-  figure~\ref{fig:repmin-AG}.
-
-  Position of the zipper is encoded using the following GADT which can be
-  generated automatically using Template Haskell:
-\begin{tabbing}\texfamily
-~~{\bfseries data}~{\itshape Position}~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}
-\\
-\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})
-\\
-\texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})
-\end{tabbing}
-  Parametrization on the type of the hole allows the code like
-  \text{\texfamily {\bfseries let}\Sp {\itshape Leaf}\Sp x\Sp =\Sp hole\Sp {\bfseries in}\Sp x} to typecheck, even though the generic zipper itself knows close
-  to nothing about the type of the hole. It might seem trivial at first, because
-  the binary tree zipper is in fact homogeneous. The ``position trick'' however
-  extends also to heterogeneous zippers which we will encounter in more advanced
-  examples.
-
-\begin{tabbing}\texfamily
-~~child~::~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Zipper}~cxt~root~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~cxt~root)
-\end{tabbing}
-  \text{\texfamily child\Sp n} moves the zipper to the \text{\texfamily n}'th child, if there is one.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}\label{sec:related-work}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}\label{sec:conclusion}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Acknowledgements}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% As per the LLNCS guidelines
-\bibliographystyle{splncs04}
-\bibliography{References}
-
-\end{document}
+  Attribute Grammars (AGs) are a declarative formalism that was proposed by
+  Knuth~\cite{Knuth68} in the late 60s and allows the implementation and
+  reasoning about programs in a modular and convenient way. A concrete AG
+  relies on a context-free grammar to define the syntax of a language, and on
+  attributes associated to

### 2018-11-21 doc/Paper.tex
Update PDF to df9deb47d64763e52c95c3526f511282576a8198

--- 2018-11-20 doc/Paper.tex
@@
 \begin{abstract}
-
   % Introduction. In one sentence, what’s the topic?
@@
   In fact, maintaining a traditional attribute grammar system is such a hard
-  effort that most such system that were proposed in the past are no longer
+  effort that most such systems that were proposed in the past are no longer
   active.
@@
   approach one then exploits the modern constructions and infrastructure that
-  are already provided by those languages and focus on the particularities of
+  are already provided by those languages and focuses on the particularities of
   the domain specific language being developed.
 
-  Functional zipper\cite{huet1997zipper} is a powerful abstraction which greatly
+  Functional zipper~\cite{huet1997zipper} is a powerful abstraction which greatly
   simplifies the implementation of traversal algorithms performing a lot of
   local updates. Functional zippers have successfully been applied to constuct
-  an attribute grammar embeddings in Haskell\cite{...}. Despite its elegance,
+  an attribute grammar embedding in Haskell~\cite{DBLP:conf/sblp/MartinsFS13,MARTINS20162}. Despite its elegance,
   this solution had a major drawback which prevented its use in real-world
@@
   which severely hurt performance. Recently, this flaw has been
-  eliminated\cite{...} and replaced with a different one: the approach became
+  eliminated~\cite{FERNANDES2018} and replaced with a different one: the approach became
   intrusive, i.e. to benefit from the embedding user-defined data structures
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Functional Zippers}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Attribute Grammars}

### 2018-11-21 build-pdf.sh
Installation instructions

@@ -1,0 +1,46 @@
+#!/bin/bash
+
+#
+# This file can be used as an example of how to build the PDF locally.
+#
+
+set -e
+
+_build_using_latexrun()
+{
+	latexrun -O "$1" "$2"
+}
+
+_build_using_pdflatex()
+{
+	declare -r output_dir="$1"
+	declare -r src_file="$2"
+	declare -r basename="${src_file%.tex}"
+	pdflatex -interaction nonstopmode -output-directory "${output_dir}" "${src_file}"
+	pdflatex -interaction nonstopmode -output-directory "${output_dir}" "${src_file}"
+	bibtex "${output_dir}/${basename}.aux"
+	pdflatex -interaction nonstopmode -output-directory "${output_dir}" "${src_file}"
+	pdflatex -interaction nonstopmode -output-directory "${output_dir}" "${src_file}"
+	mv -v "${output_dir}/${basename}.pdf" "."
+}
+
+_build()
+{
+	[[ $# -eq 2 ]] || { printf "Usage: _build <output_dir> <tex_file>"; return 1; }
+	if ! which latexrun; then
+		_build_using_latexrun "$1" "$2"
+	else
+		_build_using_pdflatex "$1" "$2"
+	fi
+}
+
+
+stack --no-terminal exec --cwd paper -- lhs2TeX Paper.lhs --tt --set=abstract_only -o Paper.tex
+pushd paper/
+mkdir -pv _tex_build/
+_build _tex_build Paper.tex
+popd
+rm -v paper/Paper.tex
+mv -v paper/Paper.pdf .
+rm -rv paper/_tex_build
+

### 2018-11-21 doc/Paper.tex
Update PDF to 787cbdae025f631eb54b7ca7c6e3787074bed149

--- 2018-11-21 doc/Paper.tex
@@
   relies on a context-free grammar to define the syntax of a language, and on
-  attributes associated to
+  attributes associated to the productions of the grammar to define the
+  semantics of that language.
+  %while adding \emph{attributes} to it so that it is also possible to define
+  %its semantics.
+  AGs have been used in practice to specify real programming languages, like
+  for example Haskell \cite{DijkstraFS09}, as well as powerful pretty printing
+  algorithms \cite{SPS99}, deforestation techniques \cite{joao07pepm} and
+  powerful type systems \cite{MiddelkoopDS10}.
+
+  When programming with AGs, modularity is achieved due to the possibility of
+  defining and using different aspects of computations as separate attributes.
+  Attributes are distinct computation units, typically quite simple and modular,
+  that can be combined into elaborated solutions to complex programming
+  problems. They can also be analyzed, debugged and maintained independently
+  which eases program development and evolution.
+
+  AGs have proven to be particularly useful to specify computations over
+  trees: given one tree, several AG systems such
+  as~\cite{syngen,uuag,lrc,silver} take specifications of which values, or
+  attributes, need to be computed on the tree and perform these computations.
+  The design and coding efforts put into the creation, improvement and
+  maintenance of these AG systems, however, is tremendous, which often is an
+  obstacle to achieving the success they deserve.
+
+  An increasingly popular alternative approach to the use of AGs relies on
+  embedding them as first class citizens of general purpose programming
+  languages~\cite{Oege00,DBLP:conf/sblp/MartinsFS13,erlangAGs,kiama,doaitse09icfp,balestrieri}.
+  This avoids the burden of implementing a totally new language and associated
+  system by hosting it in state-of-the-art programming languages. Following this
+  approach one then exploits the modern constructions and infrastructure that
+  are already provided by those languages and focuses on the particularities of
+  the domain specific language being developed.
+
+  Functional zipper~\cite{huet1997zipper} is a powerful abstraction which greatly
+  simplifies the implementation of traversal algorithms performing a lot of
+  local updates. Functional zippers have successfully been applied to constuct
+  an attribute grammar embedding in Haskell~\cite{DBLP:conf/sblp/MartinsFS13,MARTINS20162}. Despite its elegance,
+  this solution had a major drawback which prevented its use in real-world
+  applications: attributes were not cached, but rather repeatedly recomputed
+  which severely hurt performance. Recently, this flaw has been
+  eliminated~\cite{FERNANDES2018} and replaced with a different one: the approach became
+  intrusive, i.e. to benefit from the embedding user-defined data structures
+  have to be adjusted.
+
+  In this paper we present an alternative mechanism to cache attributes based on
+  a self-organising infinite grid. This graph is laid on top of the user-defined
+  algebraic data type and mirrors its structure. The used-defined data type
+  remains untouched. The embedding is then based on two coherent zippers (rather
+  than one) traversing the data structures in parallel. On top of being
+  non-intrusive our solution is completely type-safe. Modern Haskell extensions
+  such as \texttt{ConstraintKinds} allow us to propagate constraints down in the
+  ADT completely eliminating run-time type casts present in the previous
+  versions. Another side benefit of using modern Haskell is a cleaner syntax
+  with less code being generated with Template Haskell.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% As per the LLNCS guidelines
+\bibliographystyle{splncs04}
+\bibliography{References}
+
+\end{document}

### 2018-11-22 build-pdf.sh
A typo fix

@@ -27,7 +27,7 @@ _build_using_pdflatex()
 _build()
 {
 	[[ $# -eq 2 ]] || { printf "Usage: _build <output_dir> <tex_file>"; return 1; }
-	if ! which latexrun; then
+	if which latexrun; then
 		_build_using_latexrun "$1" "$2"
 	else
 		_build_using_pdflatex "$1" "$2"

### 2018-11-22 doc/Paper.tex
Update PDF to ad631252d68a67fd9dd12e877910ae96616e562e

--- 2018-11-21 doc/Paper.tex
@@
   Universidade do Minho, Portugal, \email{saraiva@di.uminho.pt} \and
-  Radboud University, The Netherlands, \email{twesterhout@student.ru.nl}%
+  Radboud University, The Netherlands, \email{t.westerhout@student.ru.nl}%
 }
@@
 
+  % 1) Introduction. In one sentence, what’s the topic?
+  % 2) State the problem you tackle
+  % 3) Summarize (in one sentence) why nobody else has adequately answered the
+  %    research question yet.
+  % 4) Explain, in one sentence, how you tackled the research question.
+  % 5) In one sentence, how did you go about doing the research that follows
+  %    from your big idea.
+  % 6) As a single sentence, what’s the key impact of your research?
 \begin{abstract}
-  % Introduction. In one sentence, what’s the topic?
-
   Attribute grammars are a powerful, declarative formalism to implement and
-  reason about programs which, by design, are conveniently modular.
-
-  % State the problem you tackle
-
-  Although a full attribute grammar compiler can be tailored to specific needs,
-  its implementation is highly non trivial, and its long term maintenance is a
-  major endeavor.
-
-  % Summarize (in one sentence) why nobody else has adequately answered the
-  % research question yet.
-
-  In fact, maintaining a traditional attribute grammar system is such a hard
-  effort that most such systems that were proposed in the past are no longer
-  active.
-
-  % Explain, in one sentence, how you tackled the research question.
-
-  Our approach to implement attribute grammars is to write them as first class
-  citizens of a modern functional programming language.
-
-  % In one sentence, how did you go about doing the research that follows from
-  % your big idea.
-
-  We improve a previous zipped-based attribute grammar embedding making it
-  non-intrusive (i.e. no changes need to be made to the user-defined data types)
-  and type-safe. On top of that, we achieve clearer syntax by using modern
-  Haskell extensions.
-
-  % As a single sentence, what’s the key impact of your research?
-
-  We believe our embedding can be employed in practice to implement elegant,
-  efficient and modular solutions to real life programming challenges.
+  reason about programs which, by design, are conveniently modular. Although a
+  full attribute grammar compiler can be tailored to specific needs, its
+  implementation is highly non-trivial, and its long-term maintenance is a major
+  endeavor. In fact, maintaining a traditional attribute grammar system is such
+  a big effort that most systems that were proposed in the past are no longer
+  active. Our approach to implementing attribute grammars is to write them as
+  first-class citizens of a modern functional programming language. We improve a
+  previous zipper-based attribute grammar embedding making it non-intrusive
+  (i.e. no changes to the user-defined data types are required) and type-safe.
+  On top of that, we achieve clearer syntax by using modern Haskell extensions.
+  We believe that our embedding can be employed in practice to implement
+  elegant, efficient, and modular solutions to real-life programming challenges.
 
@@
 
-  Attribute Grammars (AGs) are a declarative formalism that was proposed by
-  Knuth~\cite{Knuth68} in the late 60s and allows the implementation and
-  reasoning about programs in a modular and convenient way. A concrete AG
-  relies on a context-free grammar to define the syntax of a language, and on
-  attributes associated to
+  Attribute Grammars (AGs) are a declarative formalism which was proposed by
+  Knuth~\cite{Knuth68} in the late 60s and allows one to implement and reason
+  about programs in a modular and convenient way. A concrete AG relies on a
+  context-free grammar to define the syntax of a language, and on attributes
+  associated with the productions of the grammar to define the semantics of that
+  language. AGs have been used in practice to specify real programming
+  languages, like for example Haskell~\cite{DijkstraFS09}, as well as powerful
+  pretty printing algorithms~\cite{SPS99}, deforestation
+  techniques~\cite{joao07pepm}, and powerful type systems~\cite{MiddelkoopDS10}.
+
+  When programming with AGs, modularity is achieved due to the possibility of
+  defining and using different aspects of computations as separate attributes.
+  Attributes are distinct computation units, typically quite simple and modular,
+  that can be combined into elaborated solutions to complex programming
+  problems. They can also be analyzed, debugged, and maintained independently
+  which eases program development and evolution.
+
+  AGs have proven to be particularly useful to specify computations over trees:
+  given one tree, several AG systems such as~\cite{syngen,uuag,lrc,silver} take
+  specifications of which values or attributes need to be computed and perform
+  these computations. The design and coding efforts put into the creation,
+  improvement, and maintenance of these AG systems, however, is tremendous which
+  is often an obstacle to achieving the success they deserve.
+
+  An increasingly popular alternative approach to the use of AGs relies on
+  embedding them as first-class citizens of general purpose programming
+  languages~\cite{Oege00,DBLP:conf/sblp/MartinsFS13,erlangAGs,kiama,doaitse09icfp,balestrieri}.
+  This avoids the burden of implementing a totally new language and associated
+  system by hosting it in state-of-the-art programming languages. Following this
+  approach one then exploits the modern constructions and infrastructure that
+  are already provided by those languages and focuses on the particularities of
+  the domain specific language being developed.
+
+  Functional zipper~\cite{huet1997zipper} is a powerful abstraction which
+  greatly simplifies the implementation of traversal algorithms performing a lot
+  of local updates. Functional zippers have successfully been applied to
+  construct an AG embedding in
+  Haskell~\cite{DBLP:conf/sblp/MartinsFS13,MARTINS20162}. Despite its elegance,
+  this solution had a major drawback which prevented its use in real-world
+  applications: attributes were not cached, but rather repeatedly recomputed
+  which severely hurt performance. Recently, this flaw has been
+  eliminated~\cite{FERNANDES2018} and replaced with a different one: the
+  approach became intrusive, i.e. to benefit from the embedding user-defined
+  data structures have to be adjusted.
+
+  In this paper we present an alternative mechanism to cache attributes based on
+  a self-organising infinite grid. This graph is laid on top of the user-defined
+  algebraic data type (ADT) and mirrors its structure. The used-defined data type
+  itself remains untouched. The embedding is then based on two (rather than one)
+  coherent zippers traversing the data structures in parallel. On top of being
+  non-intrusive our solution is completely type-safe. Modern Haskell extensions
+  such as \texttt{ConstraintKinds} allow us to propagate constraints down in the
+  ADT completely eliminating run-time type casts present in the previous
+  versions. Another side benefit of using modern Haskell features is a cleaner
+  syntax with less code being generated by means of Template Haskell.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% As per the LLNCS guidelines
+\bibliographystyle{splncs04}
+\bibliography{References}
+
+\end{document}

### 2018-12-10 doc/Paper.tex
Update PDF to a8bf478392329f3e8b9f53803fb39d81d3672f06

@@
 \title{Modern type-safe embedding of attribute grammars}
-\subtitle{An exercise with functional zippers in Haskell}
+\subtitle{An exercise with functional zippers in Haskell\\--- Extended Abstract ---}
 
@@
   % 6) As a single sentence, what’s the key impact of your research?
-\begin{abstract}
+\section*{Abstract}
   Attribute grammars are a powerful, declarative formalism to implement and
@@
 }
-\end{abstract}
+%\end{abstract}
 

### 2018-12-11 doc/Paper.tex
Update PDF to 1ae0aa252551bffbf50a358a31fc90c1d162aab5

--- 2018-11-22 doc/Paper.tex
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Functional Zippers}
+  Zipper is a data structure commonly used in functional programming for
+  traversal with fast local updates. The zipper data structure was originally
+  conceived by Huet\cite{huet1997zipper} in the context of trees. We will,
+  however, first consider a simpler problem: a bidirectional list traversal.
+
+  \paragraph{Lists} Suppose that we would like to update a list at a specific
+  position:
+\begin{tabbing}\texfamily
+~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]\\
+\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0\\
+\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j\\
+\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after\\
+\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)\\
+\texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
+\end{tabbing}
+  Here \text{\texfamily modify} takes an update action \text{\texfamily f}\footnote{\text{\texfamily f} returns a list rather
+  than a single element to prevent curious readers from suggesting to use a
+  boxed array instead of a list.}, an index \text{\texfamily i}, and a list \text{\texfamily xs} and returns a
+  new list with the \text{\texfamily i}'th element replaced with the result of \text{\texfamily f}.%
+  First, we ``unpack'' the list into $\text{\texfamily before\Sp =\Sp [}\ttsub{\text{\texfamily xs}}{\text{\texfamily 0}},\dots
+  \ttsub{\text{\texfamily xs}}{\text{\texfamily i-1}}\text{\texfamily ]}$, $\text{\texfamily x\Sp =\Sp }\ttsub{\text{\texfamily xs}}{\text{\texfamily i}}$, and
+  $\text{\texfamily after\Sp =\Sp [}\ttsub{\text{\texfamily xs}}{\text{\texfamily i+1}},\dots\text{\texfamily ]}$. Then replace \text{\texfamily x} by \text{\texfamily f(x)} and
+  finally ``pack'' the result back into a single list. If we do a lot of
+  updates, we end up unpacking and packing the list over and over again -- very
+  time-consuming for long lists. We would thus like to benefit from fusion
+  without explicitly working with the unpacked representation as it is
+  bug-prone. A list zipper provides a way to achieve this.
+
+  A zipper consists of a focus (alternatively called a hole) and surrounding
+  context\footnote{`\text{\texfamily !}' in front of the type of \text{\texfamily \char95 cxt} is a
+  strictness annotation enabled by the \text{\ttfamily BangPatterns} extension. Context can be
+  seen as a path from the to the current focus and is always finite, i.e. there
+  is no reason for it to be lazy.}:
+\begin{tabbing}\texfamily
+~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 \\
+\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
+\end{tabbing}
+  where the \text{\texfamily {\itshape Context}} keeps track of elements to the left (\text{\texfamily before} in
+  \text{\texfamily modify}) and to the right (\text{\texfamily after} in \text{\texfamily modify}) of the focus. We can now
+  define movements:
+\begin{tabbing}\texfamily
+~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}\\
+\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$\\
+\texfamily ~~~~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)\\
+\texfamily \\
+\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}\\
+\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$\\
+\texfamily ~~~~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
+\end{tabbing}
+  and functions for entering and leaving the zipper:
+\begin{tabbing}\texfamily
+~~enter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~enter~[]~~~~~~~=~{\itshape Nothing}\\
+\texfamily ~~enter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)\\
+\texfamily \\
+\texfamily ~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]\\
+\texfamily ~~leave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
+\end{tabbing}
+
+  Finally, we define a local version of our \text{\texfamily modify} function (\TODO{Boy, is
+  this function ugly...})
+\begin{tabbing}\texfamily
+~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~modify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}\\
+\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))\\
+\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}\\
+\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')\\
+\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}\\
+\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)\\
+\texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
+\end{tabbing}
+  using which we can perform multiple updates efficiently and with minimal code
+  bloat\footnote{%
+  Operator \text{\ttfamily \char62{}\char61{}\char62{}} comes from \text{\texfamily {\itshape {\itshape Control}.Monad}} module in \text{\ttfamily base}
+  and has the following signature:%
+\begin{tabbing}\texfamily
+~~(>=>)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
+\end{tabbing}
+%
+  }:
+\begin{tabbing}\texfamily
+~~modifyExample~::~{\itshape IO}~()\\
+\texfamily ~~modifyExample~=~print~\$\\
+\texfamily ~~~~enter~@{\itshape Int}~>=>~right\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~right\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(const~[])\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~(+1))\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~left\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~negate)\\
+\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~leave~\$~[1,~2,~3,~4,~5]
+\end{tabbing}
+
+  \paragraph{Trees} Consider now a binary tree data structure:
+\begin{tabbing}\texfamily
+~~{\bfseries data}~{\itshape Tree}~a~=~{\itshape Leaf}~!a~|~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
+\end{tabbing}
+  A binary tree zipper is slightly more insteresting than the list zipper,
+  because we can move up and down the tree as well as left and right. The zipper
+  again consists of a hole (a subtree we are focused on) and its surrounding
+  context (a path from the hole to the root of the tree):
+\begin{tabbing}\texfamily
+~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~{\itshape Tree}~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 \\
+\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Top}\\
+\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)\\
+\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Right}~({\itshape Tree}~a)~!({\itshape Context}~a)
+\end{tabbing}
+  To move the zipper down, we ``unpack'' the current hole:
+\begin{tabbing}\texfamily
+~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~~~\char95 )~~~=~{\itshape Nothing}\\
+\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Right}~l~cxt)
+\end{tabbing}
+  \text{\texfamily {\itshape Context}} stores everything we need to reconstruct the hole, and \text{\texfamily up} does
+  exactly that:
+\begin{tabbing}\texfamily
+~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)\\
+\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}\\
+\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt\\
+\texfamily ~~up~({\itshape Zipper}~r~({\itshape Right}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
+\end{tabbing}
+  Implementations of \text{\texfamily left}, \text{\texfamily right}, and \text{\texfamily enter} are very similar to the
+  list zipper case and are left as an exercise for the reader. \text{\texfamily leave} differs
+  slightly in that we now move all the way up rather than left:
+\begin{tabbing}\texfamily
+~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a\\
+\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}\\
+\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'\\
+\texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
+\end{tabbing}
+
+  \paragraph{Generic Zipper} The list and binary tree zippers we have considered
+  so far are homogeneous zippers: the type of the focus does not change upon zipper
+  movement. Such a zipper can be a very useful abstraction. For example, a
+  well-known window manager XMonad\cite{xmonad} uses a rose tree zipper to track
+  the window under focus. For other tasks, however, one might need to traverse
+  heterogeneous structures. A zipper that can accomodate such needs is usually
+  called a \textit{generic zipper} as it relies only on the generic structure of
+  Algebraic Data Types (ADTs). One can view an ADT as an Abstract Syntax Tree
+  (AST) where each node is a Haskell constructor rather than a syntax construct.
+
+  The generic zipper we will use is very similar to the one presented
+  in\cite{adams2010syz}. The most common technique in Haskell for supporting
+  heterogeneous types is Existential Quantification. However, not every type can
+  act as a hole. To support moving down the tree, we need the hole to be
+  \textit{dissectible}, i.e. we would like to be able to dissect the value into
+  the constructor and its arguments. Even though \text{\texfamily {\itshape Data}.{\itshape Data}.gfoldl} allows us to
+  achieve this, we define out own typeclass which additionally allows us to
+  propagate down arbitrary constraints:
+\begin{tabbing}\texfamily
+~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}\\
+\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a\\
+\texfamily \\
+\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}\\
+\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b\\
+\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)\\
+\texfamily ~~~~~~~~~~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
+\end{tabbing}
+  For example, here is how we can make \text{\texfamily {\itshape Tree}} an instance of \text{\texfamily {\itshape Disssectible}}
+\begin{tabbing}\texfamily
+~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}\\
+\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r\\
+\texfamily ~~~~dissect~x~=~{\itshape LOne}~x
+\end{tabbing}
+  We can unpack a \text{\texfamily {\itshape Fork}} and the zipper will thus be able to go down. \text{\texfamily {\itshape Leaf}}s,
+  however, are left untouched and trying to go down from a \text{\texfamily {\itshape Leaf}} will return
+  \text{\texfamily {\itshape Nothing}}.
+
+  To allow the zipper to move left and right, we need a means to encode
+  arguments to the right of the hole. Following Adams et al, we define a GADT
+  representing constructor arguments to the right of the hole:
+\begin{tabbing}\texfamily
+~~{\bfseries data}~{\itshape Right}~c~provides~r~{\bfseries where}\\
+\texfamily ~~~~{\itshape RNil}~~::~{\itshape Right}~c~r~r\\
+\texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)\\
+\texfamily ~~~~~~~~~~=>~b~\WithMath{->}{\rightarrow}~{\itshape Right}~c~provides~r~\WithMath{->}{\rightarrow}~{\itshape Right}~c~(b~\WithMath{->}{\rightarrow}~provides)~r
+\end{tabbing}
+  For example, for a tuple \text{\texfamily ({\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int})}, we can have
+\begin{tabbing}\texfamily
+~~lefts~=~{\itshape LOne}~(,,,,,)~`{\itshape LCons}`~1~`{\itshape LCons}`~2~`{\itshape LCons}`~3\\
+\texfamily ~~hole~=~4\\
+\texfamily ~~rights~=~5~`{\itshape RCons}`~6~`{\itshape RCons}`~{\itshape RNil}
+\end{tabbing}
+  generalising this a little, we arrive at:
+\begin{tabbing}\texfamily
+~~{\bfseries data}~{\itshape LocalContext}~c~hole~rights~parent~=\\
+\texfamily ~~~~{\itshape LocalContext}~!({\itshape Left}~c~(hole~\WithMath{->}{\rightarrow}~rights))~!({\itshape Right}~c~rights~parent)\\
+\texfamily \\
+\texfamily ~~{\bfseries data}~{\itshape Context}~::~({\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}\\
+\texfamily ~~~~{\itshape RootContext}~::~\char'24~c~root.~{\itshape Context}~c~root~root\\
+\texfamily ~~~~(:>)~::~\char'24~c~parent~root~hole~rights.~(c~parent,~{\itshape Dissectible}~c~parent)\\
+\texfamily ~~~~~~~~~=>~!({\itshape Context}~c~parent~root)\\
+\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~!({\itshape LocalContext}~c~hole~rights~parent)\\
+\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Context}~c~hole~root
+\end{tabbing}
+  And just like before the \text{\texfamily {\itshape Zipper}} is a product of the hole and context:
+\begin{tabbing}\texfamily
+~~{\bfseries data}~{\itshape Zipper}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(root~::~{\itshape Type})~=\\
+\texfamily ~~~~\char'24~hole.~(c~hole,~{\itshape Dissectible}~c~hole)~=>\\
+\texfamily ~~~~~~{\itshape Zipper}~\char123 ~\char95 zHole~::~!hole\\
+\texfamily ~~~~~~~~~~~~~,~\char95 zCxt~~::~!({\itshape Context}~c~hole~root)\\
+\texfamily ~~~~~~~~~~~~~\char125 
+\end{tabbing}
+
+  Implementation of movements is quite straightforward and is left out. Please,
+  refer to (\TODO{github repo}) for complete code. We now consider a rather
+  interesting application of generic zipper: embedding of attribute grammars.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Attribute Grammars}
+  Attribute grammars (AGs) are an extension of context-free grammars that allow
+  to specify context-sensitive syntax as well as the semantics. AGs achieve it
+  by associating a set of attributes with each grammar symbol. These attributes
+  are defined using evaluation rules assiciated with production rules of the
+  context-free grammar.
+
+  Attributes are then usually divided into two disjoint sets: synthesized
+  attributes and the inherited attributes. Such distinction is required for the
+  construction of a dependency graph. It is then used for specification of the
+  evaluation order and detection of circularity. In the zipper-based embedding of
+  attribute grammars we make no use of a dependency graph and thus do not divide
+  attributes into classes.
+
+  Let us consider the repmin problem as an example of a problem that requires
+  multiple traversals:
+  \begin{displayquote}
+    Given a tree of integers, replace every integer with the
+    minimum integer in the tree, in one pass.
+  \end{displayquote}
+  The classical solution is the following circular program:
+\begin{tabbing}\texfamily
+~~repmin~::~{\itshape Tree}~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Tree}~{\itshape Int}\\
+\texfamily ~~repmin~t~=~t'\\
+\texfamily ~~~~{\bfseries where}~(t',~m')~=~go~t~m'\\
+\texfamily ~~~~~~~~~~go~({\itshape Leaf}~x~~~~)~m~=~({\itshape Leaf}~m,~x)\\
+\texfamily ~~~~~~~~~~go~({\itshape Fork}~xs~ys)~m~=~({\itshape Fork}~xs'~ys',~min~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$}~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})\\
+\texfamily ~~~~~~~~~~~~{\bfseries where}~(xs',~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$})~=~go~xs~m\\
+\texfamily ~~~~~~~~~~~~~~~~~~(ys',~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})~=~go~ys~m
+\end{tabbing}
+
+  Although quite elegant, the code lacks modularity and is very difficult to
+  reason about. Attribute Grammars provide a more modular approach. Viera et
+  al\cite{viera2009agsfly} identified three steps for solving repmin: computing
+  the minimal value, passing it down from the root to the leaves, and
+  constructing the resulting tree. We can associate each step with an
+  attribute\cite{swierstra1998designing}:
+  \begin{itemize}
+  \item A synthesized attribute \text{\texfamily localMin\Sp ::\Sp {\itshape Int}} represents the minimum value
+        of a subtree. Computing the minimal value thus corresponds to evaluation
+        of the \text{\texfamily localMin} attribute for the root tree.
+  \item An inherited attribute \text{\texfamily globalMin\Sp ::\Sp {\itshape Int}} is used to pass down the
+        minimal value.
+  \item Finally, a synthesized attribute \text{\texfamily updated\Sp ::\Sp {\itshape Tree}\Sp {\itshape Int}} is the subtree
+        with leaf values replaced by values of their \text{\texfamily globalMin} attributes. The
+        solution is thus the value of \text{\texfamily updated} attribute for the root tree.
+  \end{itemize}
+  The obtained AG is presented in figure~\ref{fig:repmin-AG}. \TODO{Do we
+  actually need to explain the algorithms here? It seems a little bit childish
+  to explain how to compute the minimum of a binary tree... If we absolutely
+  have to explain stuff, maybe just put it in the caption.}
+
+\begin{figure}
+\centering
+\fbox{\parbox{\linewidth}{%
+\begin{tabbing}\texfamily
+~~SYN~{\itshape Tree}~{\itshape Int}~[localMin~:~{\itshape Int}]\\
+\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.localMin~=~@value\\
+\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~min~@left.localMin\\
+\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.localMin\\
+\texfamily \\
+\texfamily ~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]\\
+\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin\\
+\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated\\
+\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.updated\\
+\texfamily \\
+\texfamily \\
+\texfamily ~~INH~{\itshape Tree}~{\itshape Int}~[~globalMin~:~{\itshape Int}~]\\
+\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Fork}~left.globalMin~~=~@lhs.globalMin\\
+\texfamily ~~~~~~~~~~~~~~~~~~~~~~right.globalMin~=~@lhs.globalMin\\
+\texfamily \\
+\texfamily ~~DATA~{\itshape Root}~|~{\itshape Root}~tree~:~{\itshape Tree}~{\itshape Int}\\
+\texfamily ~~SEM~~{\itshape Root}~|~{\itshape Root}~tree.globalMin~=~@tree.localMin
+\end{tabbing}
+%
+}}
+\caption{%
+  Attribute grammar for repmin. The syntax is closely mirrors the one used
+  in\cite{swierstra1998designing}. \text{\texfamily SYN} and \text{\texfamily INH} introduce synthesized
+  and inherited attributes respectively. \text{\texfamily SEM} is used for defining semantic
+  rules. A new data type \text{\texfamily {\itshape Root}} is introduced as it is common in the AG setting
+  to ``connect'' \text{\texfamily localMin} with \text{\texfamily globalMin}.
+}
+\label{fig:repmin-AG}
+\end{figure}
+
+  We now move on to embed this attribute grammar into Haskell. Semantic rules
+  simply become functions, which, given a zipper, return values of the
+  attributes. For example,
+\begin{tabbing}\texfamily
+~~localMin~::~{\itshape Zipper}~({\itshape WhereAmI}~{\itshape Position})~({\itshape Tree}~{\itshape Int})~\WithMath{->}{\rightarrow}~{\itshape Int}\\
+\texfamily ~~localMin~z@({\itshape Zipper}~hole~\char95 )~=~{\bfseries case}~whereami~hole~{\bfseries of}\\
+\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Leaf}~x~=~hole~{\bfseries in}~x\\
+\texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~\WithMath{->}{\rightarrow}~{\bfseries let}~{\itshape Just}~l~=~child~0~z;~{\itshape Just}~r~=~child~1~z\\
+\texfamily ~~~~~~~~~~~~~~{\bfseries in}~min~(localMin~l)~(localMin~r)
+\end{tabbing}
+
+  Apart from the type signature, the code is pretty straightforward and closely
+  mirrors the AG we defined earlier. \text{\texfamily whereami} function allows us to ``look
+  around'' and returns the position of the zipper. Thus the \text{\texfamily {\bfseries case}} corresponds
+  to the pattern matches on the left of the vertical bars on
+  figure~\ref{fig:repmin-AG}.
+
+  Position of the zipper is encoded using the following GADT which can be
+  generated automatically using Template Haskell:
+\begin{tabbing}\texfamily
+~~{\bfseries data}~{\itshape Position}~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}\\
+\texfamily ~~~~\parbox{\widthof{CLeaf}}{\textit{C}$_\text{\textit{Leaf}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})\\
+\texfamily ~~~~\parbox{\widthof{CFork}}{\textit{C}$_\text{\textit{Fork}}$}~::~{\itshape Position}~({\itshape Tree}~{\itshape Int})
+\end{tabbing}
+  Parametrization on the type of the hole allows the code like
+  \text{\texfamily {\bfseries let}\Sp {\itshape Leaf}\Sp x\Sp =\Sp hole\Sp {\bfseries in}\Sp x} to typecheck, even though the generic zipper itself knows close
+  to nothing about the type of the hole. It might seem trivial at first, because
+  the binary tree zipper is in fact homogeneous. The ``position trick'' however
+  extends also to heterogeneous zippers which we will encounter in more advanced
+  examples.
+
+\begin{tabbing}\texfamily
+~~child~::~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Zipper}~cxt~root~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~cxt~root)
+\end{tabbing}
+  \text{\texfamily child\Sp n} moves the zipper to the \text{\texfamily n}'th child, if there is one.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Related Work}\label{sec:related-work}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}\label{sec:conclusion}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Acknowledgements}
 

### 2018-12-21 doc/Paper.tex
Update PDF to 65e6608afa4f794444349fe633d97c4bba31428a

@@
 \usepackage{comment}
+\usepackage[inline,shortlabels]{enumitem}
 \usepackage{float}
@@
 
+\newcommand{\Let}{{\sf LET}}
+
 \begin{document}
@@
   syntax with less code being generated by means of Template Haskell.
+
+\section{Running example}
+
+\newcommand{\withSub}[2]{{\parbox{\widthof{#1#2}}{#1$_\text{#2}$}}}
+\newcommand{\itWithSub}[2]{{\parbox{\widthof{#1#2}}{\textit{#1}$_\text{#2}$}}}
+
+In order to illustrate the convenience of our embedding, we will consider the
+problem of analyzing the semantic correctness of certain functional programming
+excerpts. In particular, we will focus on \text{\texfamily {\bfseries let}} expressions that are common in
+functional languages like Haskell~\cite{}, ML~\cite{}, or Scala~\cite{}.
+Throughout the paper, we shall refer to these expressions as programs in the
+\Let\ language.
+
+A program in the \Let\ language consists of instruction blocks, where each
+instruction either:
+\begin{enumerate*}[i)]
+  \item declares a variable or
+  \item defines a nested block.
+\end{enumerate*}
+When declaring a variable, the programmer must also assigned a value to it; such
+value may be a constant, the value of a variable or an expression on these
+elements. Ultimately, each block defines a value as an expression on variables
+defined in it.
+
+Examples of programs in \Let\ are given next.
+
+\begin{tabbing}\texfamily
+\withSub{program}{1}~=~{\bfseries let}~x~=~4~{\bfseries in}~x\\
+\texfamily \\
+\texfamily \withSub{program}{2}~=~{\bfseries let}~x~=~4~{\bfseries in}~x~+~3\\
+\texfamily \\
+\texfamily \withSub{program}{3}~=\\
+\texfamily ~~{\bfseries let}~x~=~4\\
+\texfamily ~~~~~~y~=~{\bfseries let}~w~=~2~{\bfseries in}~x~+~w\\
+\texfamily ~~{\bfseries in}~~x~+~y
+\end{tabbing}
+
+The values associated with \text{\texfamily \withSub{program}{1}} and \text{\texfamily \withSub{program}{2}} are, quite
+straightforwardly, \text{\texfamily 4} and \text{\texfamily 7}, respectively. As of \text{\texfamily \withSub{program}{3}}, its value of
+\text{\texfamily 10} is calculated adding the value of \text{\texfamily x}, which is \text{\texfamily 4} and the value of \text{\texfamily y},
+i.e., \text{\texfamily 6}. The value of \text{\texfamily y} is obtained by adding the value of \text{\texfamily w}, which is
+\text{\texfamily 2}, and the value of \text{\texfamily x}, which, again, is \text{\texfamily 4}.
+
+Our goal is to implement a semantic analyzer that deals with the scope rules of
+the \Let\ language. These rules are quite natural and can be described as:
+\begin{enumerate}
+  \item if a variable is used in a block, it must be declared in that same block
+  or in an outer one. The declaration of a variable, however, may occur after its
+  use;
+  \item a variable identifier may be declared at most once within the same block.
+  In an inner block, declaring a variable that has already been declared in an
+  outer one is allowed: the identifier in the local scope hides the definition of
+  the same identifier in the global one.
+\end{enumerate}
+
+Let us now consider a more complex \Let\ program:
+
+\begin{tabbing}\texfamily
+\withSub{program}{e}~=\\
+\texfamily ~~{\bfseries let}~x~=~y\\
+\texfamily ~~~~~~a~=~{\bfseries let}~y~=~4~{\bfseries in}~y~+~w\\
+\texfamily ~~~~~~x~=~5\\
+\texfamily ~~~~~~y~=~6\\
+\texfamily ~~{\bfseries in}~~x~+~a
+\end{tabbing}
+
+According to the scope rules that we have just defined, \text{\texfamily \withSub{program}{e}} contains two
+errors:
+\begin{enumerate*}[1)]
+  \item at the outer block, variable \text{\texfamily x} has been declared twice and
+  \item at the inner block, the use of variable \text{\texfamily w} has no binding occurrence at
+        all.
+\end{enumerate*}
+Notice that \text{\texfamily y} has been declared at both the inner and the outer levels, which
+in itself is not a problem (the inner declaration hides the outer one).
+
+Programs such as the ones we have presented describe the basic block-structure
+found in many languages, with the peculiarity that variables can be used before
+they are defined.
+
+We aim to implement a program that analyses \Let\ programs and computes a list
+containing the identifiers which do not obey the scope rules of the language. In
+order to facilitate the debugging phase, we require that the list of invalid
+identifiers follows the sequential structure of the program. That is to say,
+e.g., that the semantic meaning of processing \text{\texfamily \withSub{program}{e}} is \text{\texfamily [w,\Sp x]}: the use
+of the undeclared variable \text{\texfamily w} occurs in line 3 whereas the duplicate
+declaration of \text{\texfamily x} occurs in line 4.
+
+Since variables can be used before they are declared, a natural way to implement
+such an analysis is to traverse a \Let\ program twice: once to accumulate the
+declarations of identifiers (at each block), and again to check the uses of
+identifiers against such declarations. The uniqueness of declarations can
+already be detected in the first traversal: for each newly encountered
+declaration, it is possible to check whether its identifier has already been
+declared (in the same level).
+
+In order to implement such program, we first need a representation for programs
+in the \Let\ language. For this, we may use the following Haskell data-types:
+
+\begin{tabbing}\texfamily
+{\bfseries type}~{\itshape Var}~~~~=~{\itshape String}\\
+\texfamily {\bfseries data}~{\itshape Let}~~~~=~{\itshape Let}~{\itshape Decls}~{\itshape Expr}\\
+\texfamily {\bfseries data}~{\itshape Decls}~~=~~{\itshape Empty}\\
+\texfamily ~~~~~~~~~~~~|~~{\itshape Cons}~~~~{\itshape Var}~{\itshape Expr}~~{\itshape Decls}\\
+\texfamily ~~~~~~~~~~~~|~~{\itshape Nested}~~{\itshape Var}~{\itshape Let}~~~{\itshape Decls}\\
+\texfamily {\bfseries data}~{\itshape Expr}~~~=~~{\itshape Const}~~~~~{\itshape Int}\\
+\texfamily ~~~~~~~~~~~~|~~{\itshape Variable}~~{\itshape Var}\\
+\texfamily ~~~~~~~~~~~~|~~{\itshape Plus}~~~{\itshape Expr}~{\itshape Expr}\\
+\texfamily ~~~~~~~~~~~~|~~{\itshape Times}~~{\itshape Expr}~{\itshape Expr}
+\end{tabbing}
+
+In this representation, \text{\texfamily \withSub{program}{e}} above is defined as:
+
+\begin{tabbing}\texfamily
+\withSub{program}{e}~=~{\itshape Let}\\
+\texfamily ~~({\itshape Cons}\\
+\texfamily ~~~~\char34 x\char34 \\
+\texfamily ~~~~({\itshape Variable}~\char34 y\char34 )\\
+\texfamily ~~~~({\itshape Nested}\\
+\texfamily ~~~~~~\char34 a\char34 \\
+\texfamily ~~~~~~({\itshape Let}~({\itshape Cons}~\char34 y\char34 ~({\itshape Const}~4)~{\itshape Empty})~({\itshape Plus}~({\itshape Variable}~\char34 y\char34 )~({\itshape Variable}~\char34 w\char34 )))\\
+\texfamily ~~~~~~({\itshape Cons}~\char34 x\char34 ~({\itshape Const}~5)~({\itshape Cons}~\char34 y\char34 ~({\itshape Const}~6)~{\itshape Empty}))\\
+\texfamily ~~~~)\\
+\texfamily ~~)\\
+\texfamily ~~({\itshape Plus}~({\itshape Variable}~\char34 x\char34 )~({\itshape Variable}~\char34 a\char34 ))
+\end{tabbing}
+
+Now, we implement name analysis on an abstract \Let\ tree using a set of
+composable functions.
+
+The function that implements the first traversal described above needs to pass
+all the information that is required for the second traversal. Namely, in order
+to compute the final list of errors in the desired order, the second traversal
+needs to \textit{where} errors have occurred during the first. Also, the value
+of the nesting level must also be carried around because it is on the second
+traversal that the first traversal on nested expressions starts (the initial
+environment of an inner block is composed by the complete environment of its
+outer one).
+
+In order to make available all the information that the second traversal needs,
+the first traversal will build a structure such as:
+
+\begin{tabbing}\texfamily
+{\bfseries data}~\itWithSub{Let}{2}~~~=~\itWithSub{Let}{2}~\itWithSub{Decls}{2}~{\itshape Expr}\\
+\texfamily {\bfseries data}~\itWithSub{Decls}{2}~=~\itWithSub{Empty}{2}\\
+\texfamily ~~~~~~~~~~~~|~\itWithSub{Cons}{2}~{\itshape Errors}~{\itshape Expr}~\itWithSub{Decls}{2}\\
+\texfamily ~~~~~~~~~~~~|~\itWithSub{Nested}{2}~{\itshape Errors}~{\itshape Lev}~{\itshape Let}~\itWithSub{Decls}{2}
+\end{tabbing}
+
+We are now ready to implement the two traversal strategy that we have described.
+
+\begin{tabbing}\texfamily
+{\bfseries type}~{\itshape Errors}~~=~[{\itshape String}]\\
+\texfamily {\bfseries type}~{\itshape Lev}~~~~~=~{\itshape Int}\\
+\texfamily \\
+\texfamily semantics~::~{\itshape Let}~\WithMath{->}{\rightarrow}~{\itshape Errors}\\
+\texfamily semantics~program~=~errors\\
+\texfamily ~{\bfseries where}\\
+\texfamily ~~(\withSub{let}{2},~env)~=~\withSub{duplicate}{\textit{Let}}~program~[]~0\\
+\texfamily ~~errors~~~~~~~=~\withSub{missing}{\textit{Let}}~\withSub{let}{2}~env\\
+\texfamily \\
+\texfamily \withSub{duplicate}{\textit{Let}}~::~{\itshape Let}~\WithMath{->}{\rightarrow}~[({\itshape Var},~{\itshape Lev})]~\WithMath{->}{\rightarrow}~{\itshape Lev}~\WithMath{->}{\rightarrow}~(\itWithSub{Let}{2},~[({\itshape Var},~{\itshape Lev})])\\
+\texfamily \withSub{duplicate}{\textit{Let}}~({\itshape Let}~decls~expr)~dcli~lev~=~(\itWithSub{Let}{2}~\withSub{decls}{2}~expr,~dclo)\\
+\texfamily ~~{\bfseries where}~(\withSub{decls}{2},~dclo)~=~\withSub{duplicate}{\textit{Decls}}~decls~dcli~lev\\
+\texfamily \\
+\texfamily \withSub{duplicate}{\textit{Decls}}~::~{\itshape Decls}~\WithMath{->}{\rightarrow}~[({\itshape Var},~{\itshape Lev})]~\WithMath{->}{\rightarrow}~{\itshape Lev}~\WithMath{->}{\rightarrow}~(\itWithSub{Decls}{2},~[({\itshape Var},~{\itshape Lev})])\\
+\texfamily \withSub{duplicate}{\textit{Decls}}~{\itshape Empty}~dcli~lev~=~(\itWithSub{Empty}{2},~dcli)\\
+\texfamily \withSub{duplicate}{\textit{Decls}}~({\itshape Cons}~var~expr~decls)~dcli~lev~=\\
+\texfamily ~~(\itWithSub{Cons}{2}~\textit{error}~expr~\withSub{decls}{2},~dclo)\\
+\texfamily ~{\bfseries where}\\
+\texfamily ~~\textit{error}~~~~~~~~~~~=~{\bfseries if}~(var,~lev)~`elem`~dcli~{\bfseries then}~[var]~{\bfseries else}~[]\\
+\texfamily ~~(\withSub{decls}{2},~dclo)~=~\withSub{duplicate}{\textit{Decls}}~decls~((var,~lev)~:~dcli)~lev\\
+\texfamily \withSub{duplicate}{\textit{Decls}}~({\itshape Nested}~var~nested~decls)~dcli~lev~=\\
+\texfamily ~~(\itWithSub{Nested}{2}~\textit{error}~(lev~+~1)~nested~\withSub{decls}{2},~dclo)\\
+\texfamily ~{\bfseries where}\\
+\texfamily ~~\textit{error}~~~~~~~~~~~=~{\bfseries if}~(var,~lev)~`elem`~dcli~{\bfseries then}~[var]~{\bfseries else}~[]\\
+\texfamily ~~(\withSub{decls}{2},~dclo)~=~\withSub{duplicate}{\textit{Decls}}~decls~((var,~lev)~:~dcli)~lev\\
+\texfamily \\
+\texfamily \withSub{missing}{\textit{Let}}~::~\itWithSub{Let}{2}~\WithMath{->}{\rightarrow}~[({\itshape Var},~{\itshape Lev})]~\WithMath{->}{\rightarrow}~{\itshape Errors}\\
+\texfamily \withSub{missing}{\textit{Let}}~(\itWithSub{Let}{2}~decls~expr)~env~=~\withSub{errors}{1}~\WithMath{++}{+\!\!+}~\withSub{errors}{2}\\
+\texfamily ~{\bfseries where}\\
+\texfamily ~~\withSub{errors}{1}~=~\withSub{missing}{\textit{Decls}}~decls~env\\
+\texfamily ~~\withSub{errors}{2}~=~\withSub{missing}{\textit{Expr}}~expr~env\\
+\texfamily \\
+\texfamily \withSub{missing}{\textit{Decls}}~::~\itWithSub{Decls}{2}~\WithMath{->}{\rightarrow}~[({\itshape Var},~{\itshape Lev})]~\WithMath{->}{\rightarrow}~{\itshape Errors}\\
+\texfamily \withSub{missing}{\textit{Decls}}~(\itWithSub{Cons}{2}~\textit{error}~expr~decls)~env~=~\textit{error}~\WithMath{++}{+\!\!+}~errors\\
+\texfamily ~~{\bfseries where}~errors~=~\withSub{missing}{\textit{Expr}}~expr~env~\WithMath{++}{+\!\!+}~\withSub{missing}{\textit{Decls}}~decls~env\\
+\texfamily \withSub{missing}{\textit{Decls}}~(\itWithSub{Nested}{2}~\textit{error}~lev~nested~decls)~env~=~\textit{error}~\WithMath{++}{+\!\!+}~errors\\
+\texfamily ~{\bfseries where}\\
+\texfamily ~~(\withSub{nested}{2},~dclo)~=~\withSub{duplicate}{\textit{Let}}~nested~env~lev\\
+\texfamily ~~errors~~~~~~~~~~~=~\withSub{missing}{\textit{Let}}~\withSub{nested}{2}~dclo~\WithMath{++}{+\!\!+}~\withSub{missing}{\textit{Decls}}~decls~env\\
+\texfamily \withSub{missing}{\textit{Decls}}~\itWithSub{Empty}{2}~\char95 ~=~[]\\
+\texfamily \\
+\texfamily \withSub{missing}{\textit{Expr}}~::~{\itshape Expr}~\WithMath{->}{\rightarrow}~[({\itshape Var},~b)]~\WithMath{->}{\rightarrow}~{\itshape Errors}\\
+\texfamily \withSub{missing}{\textit{Expr}}~({\itshape Const}~\char95 )~\char95 ~=~[]\\
+\texfamily \withSub{missing}{\textit{Expr}}~({\itshape Plus}~\withSub{expr}{1}~\withSub{expr}{2})~env~=\\
+\texfamily ~~\withSub{missing}{\textit{Expr}}~\withSub{expr}{1}~env~\WithMath{++}{+\!\!+}~\withSub{missing}{\textit{Expr}}~\withSub{expr}{2}~env\\
+\texfamily \withSub{missing}{\textit{Expr}}~({\itshape Times}~\withSub{expr}{1}~\withSub{expr}{2})~env~=\\
+\texfamily ~~\withSub{missing}{\textit{Expr}}~\withSub{expr}{1}~env~\WithMath{++}{+\!\!+}~\withSub{missing}{\textit{Expr}}~\withSub{expr}{2}~env\\
+\texfamily \withSub{missing}{\textit{Expr}}~({\itshape Variable}~var)~env~=~{\bfseries if}~var~`elem`~map~fst~env~{\bfseries then}~[]~{\bfseries else}~[var]
+\end{tabbing}
+
+Notice that \text{\texfamily \withSub{duplicate}{\textit{Let}}} not only computes the total environment (using an
+initially empty accumulating parameter), but it also computes a \text{\texfamily \itWithSub{Let}{2}}
+intermediate data structure that stores, e.g., the duplicated variables detected
+during the first traversal. The second traversal starts with a call to
+\text{\texfamily \withSub{missing}{\textit{Let}}} giving that computed data structure and the accumulated environment
+as arguments. It produces the list of errors that follows the sequential
+structure of the program.
+
+In function \text{\texfamily \withSub{duplicate}{\textit{Decls}}}, for every block we compute: its environment, its
+level and its invalid identifiers. The environment defines the context where the
+block occurs. It consists of all the identifiers that are visible in the block
+(annotated with the level of the block). The level indicates the nesting depth
+of a block. Observe that we have to distinguish between the same identifier
+declared at different levels, which is valid.
+
+Finally, please note that in the second traversal of a nested expression, in
+function \text{\texfamily \withSub{missing}{\textit{Decls}}} for the constructor \text{\texfamily \itWithSub{Nested}{2}}, the program performs the
+two traversals to the body of that expression: calls \text{\texfamily \withSub{duplicate}{\textit{Let}}} and
+\text{\texfamily \withSub{missing}{\textit{Let}}}.
+
+Although the semantic analysis we have implemented for the \Let\ language is
+relatively simple, still we had to face some challenges.
+
+Indeed, scheduling computations was by no means trivial, with intermingled
+recursive functions. Also, we had to carefully design and implement intermediate
+data structures in order to convey data between traversals. These challenges are
+common to functional programming solutions to realistic programming problems.
+
+In lazy (functional) programming languages, one can avoid both the need for
+scheduling and for additional data structures by constructing circular
+programs~\cite{Bird84}. This strategy, however, compromises the much desired
+modular nature of the implementations.
+
+In our work, we seek an elegant and efficient alternative to the construction of
+functional programs.
 
